---
title: "Batch correction and data set integration"
subtitle: "03_Integration_Harmony"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Batch correction and data set integration

## Setup

### Standard packages

```{r LoadPackages, warning=FALSE}
library(here)
source(here("code", "standard_libraries.R"))
```

### Additional Packages

```{r LoadAddPackages, warning=FALSE}
#Load Packages

suppressPackageStartupMessages({
library(harmony)
library(intrinsicDimension)
#library(reticulate)
#library(bluster)
})

```


### Set Parameter

```{r Setup}
set.seed(100)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=100)
path <- here::here()
```


## Load Data

```{r LoadDataset}
## RA DATASET
sce <- readRDS(file = paste0(path,'/output/02_sce_DimensionalityReduction.rds'))
```

### Explore Dataset

```{r ShowDim}
#Dimensions of count matrix
dim(sce)
```

## Pre-Batch Correction Plots {.tabset}

### PCA

```{r PlotPCABB}
#Plot UMAP before data integration
plotReducedDim(sce, dimred="PCA", colour_by="Sample")
```

### UMAP

```{r PlotUMAPBB}
#Plot UMAP before data integration
plotReducedDim(sce, dimred="UMAP", colour_by="Sample")
```

## Batch Removal with Harmony

```{r InitializeSce}
#Initialize sce element with batch removal
sce.BatchR <- sce 
```

### Run Harmony 

```{r}
#Perform scaling normalization within each batch
meta_data <- colData(sce.BatchR) %>%
  as.data.frame() %>%
  dplyr::select(c(Sample)) #Add covariates zb protocol

sce.BatchR <- harmony::RunHarmony(sce.BatchR, group.by.vars=c("Sample"),reduction.save="harmony",verbose=FALSE) #  Add covariates zb protocol
```


### QC Plots Harmony

We use the percentage of variance lost as a diagnostic measure.Large proportions of lost variance (>10%) suggest that correction is removing genuine biological heterogeneity.
```{r}
# Percentage of variance lost 
metadata(sce.BatchR)$merge.info$lost.var

```

### Chosing the number of PCs

The advantage of using PCA is that the total amount of variance explained by the first few PCs is usually enough to capture most of the signal in the data. Therefore, we can exclude the remaining PCs without much loss of information. 

```{r ShowPC}
#PCA dim elbow estimate using 
elbow <- ceiling(as.numeric(maxLikGlobalDimEst(as.matrix(reducedDim(sce.BatchR, "harmony")), k=10)))

#PCA select nr of PC
reducedDim(sce.BatchR,"harmony_reduced") <- reducedDim(sce.BatchR,"harmony")[,seq_len(20)]
reducedDimNames(sce.BatchR)

ncol(reducedDim(sce.BatchR,"harmony"))
ncol(reducedDim(sce.BatchR,"harmony_reduced"))
```


## PCA Plots after Integration {.tabset}
### Colored by Sample
```{r}
#Plot PCA after data integration
plotReducedDim(sce.BatchR, dimred="harmony_reduced", colour_by="Sample") + ggtitle("PCA after data integration with harmony")
```
### Colored by Joint
```{r}
#Plot PCA after data integration
plotReducedDim(sce.BatchR, dimred="harmony_reduced", colour_by="Joint.Location") + ggtitle("PCA after data integration with harmony")
```
### Facet by Sample
```{r}
#Plot PCA after data integration per sample
plotReducedDim(sce.BatchR, dimred="harmony_reduced", colour_by="Sample") + ggtitle("PCA after data integration with harmony")+ facet_wrap(~sce.BatchR$Sample)
```
### Facet by Joint
```{r}
#Plot PCA after data integration per joint
plotReducedDim(sce.BatchR, dimred="harmony_reduced", colour_by="Joint.Location") + ggtitle("PCA after data integration with harmony")+ facet_wrap(~sce.BatchR$Joint.Location)

```


## UMAP Plots after Integration {.tabset}

```{r}
#Run UMAP
sce.BatchR <- runUMAP(sce.BatchR, dimred="harmony", name="harmony_UMAP")
sce.BatchR <- runUMAP(sce.BatchR, dimred="harmony_reduced", name="harmony_reduced_UMAP")

```

### Colored by Sample
```{r}
#Plot UMAP before data integration
plotReducedDim(sce.BatchR, dimred="harmony_reduced_UMAP", colour_by="Sample") + ggtitle("UMAP after data integration with harmony")
```
### Colored by Joint
```{r}
#Plot UMAP before data integration
plotReducedDim(sce.BatchR, dimred="harmony_reduced_UMAP", colour_by="Joint.Location") + ggtitle("UMAP after data integration with harmony")
```
### Facet by Sample
```{r}
#Plot UMAP after data integration per sample
plotReducedDim(sce.BatchR, dimred="harmony_reduced_UMAP", colour_by="Sample") + ggtitle("UMAP after data integration with harmony")+ facet_wrap(~sce.BatchR$Sample)

```
### Facet by Joint
```{r}
#Plot UMAP after data integration per joint
plotReducedDim(sce.BatchR, dimred="harmony_reduced_UMAP", colour_by="Joint.Location") + ggtitle("UMAP after data integration with harmony")+ facet_wrap(~sce.BatchR$Joint.Location)
```


## Save the dataset

```{r SaveData}
saveRDS(sce.BatchR, file =paste0(path,'/output/03_sce_Integration_Harmony.rds'))
```
