---
title: "Marker Genes and Celltype Annotation"
subtitle: "06_CelltypeAnnotation"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Marker Genes and Celltype Annotation

## Setup

### Load packages

```{r LoadPackages, warning=FALSE}
#Standard Packages
library(here)
source(here("code", "standard_libraries.R"))

#Additional Packages
suppressPackageStartupMessages({
library(dittoSeq)
library(pheatmap)
library(xlsx)
library(presto)
library(UCell)
library(patchwork)
})

```

### Set Parameter

```{r Setup}
set.seed(123)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=123)
path <- here::here()

#Output path
output_dir <- here("output","Annotation_output")
if(!dir.exists(output_dir)) dir.create(output_dir)

```


## Load Data

```{r LoadDataset}
## RA DATASET
sce <- readRDS(file = paste0(path,'/output/05_sce_Clustering.rds'))

```

### Change Rownames to the Symbol

```{r ChangeRownames}
#Change to genesymbol whenever possible
rowData(sce)$UniqID <-rownames(sce)
rownames(sce) <- uniquifyFeatureNames(rowData(sce)$ENSEMBL, rowData(sce)$Symbol)
data.frame(rowData(sce))
```


### Explore Dataset

```{r ShowDim}
#Dimensions of count matrix
dim(sce)
```


## Plot Clustering and QC {.tabset}

### PCA

```{r PlotClusterPCA}
plotReducedDim(sce, dimred="MNN_reduced", colour_by="cluster",text_by="cluster",point_alpha=0.8,point_size=1) + 
  ggtitle("PCA with clusters")+
  labs( x='PCA 1', y='PCA 2' )
```

### UMAP

```{r PlotClusterUMAP}
plotReducedDim(sce, dimred="MNN_UMAP_reduced", colour_by="cluster",text_by="cluster",point_alpha=0.8,point_size=1) + 
  ggtitle("UMAP with with clusters")+
  labs( x='UMAP 1', y='UMAP 2' )
```

### QC UMAPs

```{r PlotClusterQCUMAP,fig.height=10, fig.width=15}
print(Plot_QC_dimred(sce,dim="MNN_UMAP_reduced"))
```


## Identify Marker Genes

### Wilcoxon rank-sum test on Cluster

```{r IdentifyMarker}
#Wilcoxon rank-sum test
markers_wilcox <- wilcoxauc(sce, group_by='cluster', assay = 'logcounts')

```

```{r SaveTopMarker}
# Get the top 50 markers for each cluster (save in excel file)
# filter for nominally significant (p<0.05) and over-expressed (auc>0.5)
top50DEgenes <- top_markers(markers_wilcox, n = 50, padj_max = 0.05, pct_in_min = 20, auc_min = 0.5)
data.frame(top50DEgenes)


#Save top 50 Markers of clusters in excel sheet
wb = createWorkbook()
for (i in 1:max(as.integer(levels(sce$cluster)))){
  markers<-top50DEgenes[,paste0(i), drop=TRUE]  #get genes for the cluster
  tmp<-markers_wilcox[markers_wilcox$feature %in% markers & markers_wilcox$group == i,]#Create table
  #Save table in Exelsheet
  addDataFrame(tmp, sheet=createSheet(wb, paste0("Cluster ", i)), startColumn=1, row.names=FALSE,col.names = TRUE)
}
saveWorkbook(wb,paste0(output_dir, "/Cluster_Marker_Genes.xlsx"))

```

### Wilcoxon heatmap and dotplot

```{r PlotHeatmap,fig.height=10, fig.width=15}

# Get the top 3 markers for each cluster 
# filter for nominally significant (p<0.05)and over-expressed (auc>0.5)
top3DEgenes <- top_markers(markers_wilcox, n = 3, padj_max = 0.05, pct_in_min = 20, auc_min = 0.5)#auc_min = 0.5

#Show top 3 DE Genes of each cluster
data.frame(top3DEgenes)
top3DEgenes<-unlist(as.data.frame(top3DEgenes[,-1]),use.names = FALSE)

#Heatmap of top marker genes of each cluster by Wilcoxon rank-sum 

# To avoid a too big plot we only keep every 20th cell
sce.plot <- sce[unique(top3DEgenes), seq(1, dim(sce)[2], 20)]
#Heatmap
dittoHeatmap(sce.plot, vars = unique(top3DEgenes), annot.by = c("cluster"),order.by="cluster", scaled.to.max = TRUE,cluster_rows=FALSE,cluster_cols=FALSE) 

#Dotplot
dittoDotPlot(sce, vars = unique(top3DEgenes) , group.by = "label",min.color="lightgrey", max.color="blue",min.percent=0.2,max.percent=1,scale = TRUE )

rm(sce.plot,column_annotations,markers_wilcox_up)
gc()
```




## Cell module score with UCell

### Create cell signatures with marker genes

```{r MarkerGenes}
#Function to visuaise gene Expression for vector of genes

Coarse_signatures <- list(
    Immune = c("PTPRC","CORO1A","RAC2"),
    Myeloid = c("CSF3R","MS4A6A","MS4A7"),
    Stromal = c("PDGFRB","DCN","PCOLCE","ACTA2"),
    Endothelial = c("CDH5","ECSCR","CCL14")
)

Fine_signatures<- list(
    Fibroblast = c("PDGFRB","DCN","COL1A1","PRG4"),
    Smooth_muscle_cell = c("ACTA2","MYL9","DES","MYH11"),
    Endothelial_cell = c("CDH5","VWF", "PECAM1","CCL14"),
    Tcell = c("CD3E","CD3D", "CD3G","TRAC"),
    Bcell = c("CD19","MS4A1","CD79A","TNFRSF13C"),
    Neutrophil = c("CSF3R","S100A8","TREM1","NAMPT"),
    Mast_cell = c("TPSAB1","CPA3","HPGDS"),
    Macrophage = c("MARCO","LYZ","MSR1","CD14","FCGR3A"),
    Plasma = c("JSRP1","DERL3", "JCHAIN","SDC1","MZB1"),
    Dendritic_cell = c("FLT3","BTLA","CD86","CLEC9A","CCL22","CD1C")
)   
    
```


### Run UCell on sce object

#### Coarse signatures
```{r UCellCoarse}
#Coarse_signatures
sce.ucell <- ScoreSignatures_UCell(sce, features=Coarse_signatures, assay = 'counts', name=NULL, BPPARAM=bpp)
altExp(sce.ucell, 'UCell')
```

```{r UCellCoarseUMAP,fig.height=10, fig.width=15}
pll <- lapply(names(Coarse_signatures), function(x) {
  plotReducedDim(sce.ucell, dimred="MNN_UMAP_reduced", colour_by=x,by_exprs_values = "UCell",point_alpha=0.8,point_size=0.2) +
  labs( x='UMAP 1', y='UMAP 2' )+
  scale_color_viridis(option="magma")
})
wrap_plots(pll)
```

```{r UCellCoarseExpression,fig.height=10, fig.width=15}
pll <- lapply(names(Coarse_signatures), function(x) {
  plotExpression(sce.ucell, features = x, x = "cluster",exprs_values = "UCell")
})
wrap_plots(pll,ncol=2)
```


#### Fine signatures

```{r UCellFine}
#Fine_signatures
sce.ucell <- ScoreSignatures_UCell(sce, features=Fine_signatures, assay = 'counts', name=NULL, BPPARAM=bpp)
altExp(sce.ucell, 'UCell')
```

```{r UCellFineUMAP,fig.height=10, fig.width=15}
pll <- lapply(names(Fine_signatures), function(x) {
    plotReducedDim(sce.ucell, dimred="MNN_UMAP_reduced", colour_by=x,by_exprs_values = "UCell",point_alpha=0.8,point_size=0.2) +
  labs( x='UMAP 1', y='UMAP 2' )+
  scale_color_viridis(option="magma")
})
wrap_plots(pll)
```

```{r UCellFineExpression,fig.height=10, fig.width=15}
pll <- lapply(names(Fine_signatures), function(x) {
  plotExpression(sce.ucell, features = x, x = "cluster",exprs_values = "UCell")
})
wrap_plots(pll,ncol=2)
```



## Manual Celltype annotation

```{r VizGenExFunction}
#Function to visuaise gene Expression for vector of genes
VizGenEx <- function(markr){
 
  for(i in markr){
    cat(paste0("\n##### ", i, "\n"))
    gridExtra::grid.arrange(
      arrangeGrob(plotReducedDim(sce, dimred = "MNN_UMAP_reduced", colour_by = i, text_by = "cluster",point_alpha=1,point_size=0.5), 
                  plotReducedDim(sce, dimred = "MNN_UMAP_reduced", colour_by = i, order_by=i ,point_alpha=1,point_size=0.5) 
                    +scale_colour_gradient(name = i,low = "lightgrey", high = "red"), ncol = 2), 
                  plotExpression(sce, features = i, x = "cluster"),nrow = 2)
    cat(' \n')
  }
}

```


### Major cell type Annotation - Level0 {.tabset}

Identify major cell types for a Rough Annotation on high hierarchical level

#### Lymphocyte {.tabset}

```{r LymphocyteL0, results='asis'}
  
VizGenEx(Coarse_signatures$Immune)  
```

#### Myeloid {.tabset}

```{r MyeloidL0, results='asis'}

VizGenEx(Coarse_signatures$Myeloid)  

```

#### Stromal {.tabset}

```{r StromalL0, results='asis'}

VizGenEx(Coarse_signatures$Stromal)  

```

#### Endothelial {.tabset}

```{r EndothelialL0, results='asis'}
VizGenEx(Coarse_signatures$Endothelial)  

```

### Assignement - Level0 

```{r AssignementL0}
#Manual Annotation
sce$Annotation_L0 <- as.integer(sce$cluster)
unique(sce$Annotation_L0)

sce$Annotation_L0[sce$cluster %in% c(5,12,13,17,18)] <- "Lymphocyte"
sce$Annotation_L0[sce$cluster %in% c(4,7,8,9,10,15,19,20)] <- "Myeloid"
sce$Annotation_L0[sce$cluster %in% c(2,3,6,11,16)] <- "Stromal"
sce$Annotation_L0[sce$cluster %in% c(1,14)] <- "Endothelial"

```

### Finer cell type Annotation - Level1 {.tabset}

Identify cell types for a finer grained Annotation on lower hierarchical level

#### Macrophage {.tabset}

```{r MacrophageL1, results='asis'}
VizGenEx(Fine_signatures$Macrophage) 
```


#### Fibroblast {.tabset}

```{r FibroblastL1, results='asis'}
VizGenEx(Fine_signatures$Fibroblast) 
```

#### Endothelial {.tabset}

```{r EndothelialL1, results='asis'}
VizGenEx(Fine_signatures$Endothelial_cell) 

```


#### BCell {.tabset}

```{r BCellL1, results='asis'}
VizGenEx(Fine_signatures$Bcell) 
```

#### Plasma {.tabset}

```{r PlasmaL1, results='asis'}
VizGenEx(Fine_signatures$Plasma)
```

#### Tcell {.tabset}

```{r TcellL1, results='asis'} 
VizGenEx(Fine_signatures$Tcell) 
```

#### Smooth muscle cell {.tabset}

```{r SMCL1, results='asis'}
VizGenEx(Fine_signatures$Smooth_muscle_cell) 
```

#### Mastcell {.tabset}

```{r MastcellL1, results='asis'}
VizGenEx(Fine_signatures$Mast_cell) 
```

#### Neutrophil {.tabset}

```{r NeutrophilL1, results='asis'}
VizGenEx(Fine_signatures$Neutrophil) 
```

#### Dendritic cell {.tabset}

```{r DCL1, results='asis'}
VizGenEx(Fine_signatures$Dendritic_cell) 

```


#### Proliferating Cells{.tabset}

```{r ProliferatingL1, results='asis'}
#VizGenEx(Fine_signatures$) 
VizGenEx(c("MKI67","PCNA","STMN1","HMGB2")) 

```


### Assignement - Level1 

```{r AssignementL1}
#Manual Annotation
sce$Annotation_L1 <- as.integer(sce$cluster)
unique(sce$Annotation_L1)

sce$Annotation_L1[sce$cluster %in% c(2,3,6,16)] <- "Fibroblast"
sce$Annotation_L1[sce$cluster %in% c(11)] <- "Smooth muscle cell"
sce$Annotation_L1[sce$cluster %in% c(1,14)] <- "Endothelial cell"
sce$Annotation_L1[sce$cluster %in% c(12,13,17)] <- "T cell"
sce$Annotation_L1[sce$cluster %in% c(18)] <- "B cell"
sce$Annotation_L1[sce$cluster %in% c(20)] <- "Neutrophil"
sce$Annotation_L1[sce$cluster %in% c(8)] <- "Mast cell"
sce$Annotation_L1[sce$cluster %in% c(4,9,10,15,19)] <- "Macrophage"
sce$Annotation_L1[sce$cluster %in% c(5)] <- "Plasma"
sce$Annotation_L1[sce$cluster %in% c(7)] <- "Dendritic cell"

```



```{r AssignementCheck}
#Check if all clusters assigned
unique(sce$Annotation_L0)
unique(sce$Annotation_L1)
```

#### Plot Assigned Celltypes {.tabset}

##### UMAP L1

```{r PlotL1}
#Plot manual Annotation
plotReducedDim(sce, dimred="MNN_UMAP_reduced", colour_by="Annotation_L1",point_alpha=0.7,point_size=0.1)+
  scale_colour_manual( values=meta_colors$celltype_level1[seq.int(length(unique(sce$Annotation_L1)))],name = "Annotation" )+ 
  labs( x='UMAP 1', y='UMAP 2' )+
  guides(colour = guide_legend(override.aes = list(size=2)))
```

##### UMAP L0

```{r PlotL0}
#Plot manual Annotation
plotReducedDim(sce, dimred="MNN_UMAP_reduced", colour_by="Annotation_L0",point_alpha=0.7,point_size=0.1)+
  scale_colour_manual( values=meta_colors$celltype_level0[seq.int(length(unique(sce$Annotation_L0)))],name = "Annotation" )+ 
  labs( x='UMAP 1', y='UMAP 2' )+
  guides(colour = guide_legend(override.aes = list(size=2)))
```


## Set the Final Celllabels

```{r SetLabel}
## Set Celltypes to use downstream as colLabels
colLabels(sce) <- colData(sce)$Annotation_L1
sce$celltype <- colData(sce)$Annotation_L1

```

## Plot the assigned Cell labels

```{r PlotLabel}
plotReducedDim(sce, dimred="MNN_UMAP_reduced", colour_by="celltype",point_alpha=0.7,point_size=0.1)+
  scale_colour_manual( values=meta_colors$celltype_level1[seq.int(length(unique(sce$celltype)))],name = "Annotation" )+ 
  labs( x='UMAP 1', y='UMAP 2' )+
  guides(colour = guide_legend(override.aes = list(size=2)))
```


## Plot Selected Markergenes 

```{r PlotLabelHeatmap}
#Heatmap of Marker Gened used for Annotation
#Endo "VWF", "CDH5"
#Macrophage "CD14","FCGR3A"
#Bcell "CD19","MS4A1","CD79A"
#Plasma "SDC1","MZB1"
#Tcell "CD3E", "CD3G"
#Mast "KIT","IL1RL1","MS4A2" "TPSAB1","CPA3"
#Fibroblast "PDGFRB","DCN","COL1A1"
#Neutrophil "CSF3R","S100A8","TREM1"
#DC "FLT3","BTLA","CD86","CLEC9A","CCL22","CD1C"
#SMC

markers <- c("PDGFRB","COL1A1","ACTA2","VWF", "CDH5","CD14","FCGR3A","CSF3R","S100A8","FLT3","CD1C","TPSAB1","CPA3","KIT","MS4A2","CD19","MS4A1","CD79A","SDC1","MZB1","CD3E", "CD3G")



# To avoid a too big plot we only keep every 20th cell
sce.plot <- sce[unique(markers), seq(1, dim(sce)[2], 20)]
dittoHeatmap(sce.plot, vars = unique(markers), annot.by = c("label"),scaled.to.max = TRUE,cluster_rows=FALSE,cluster_cols=FALSE)


#Dotplot
dittoDotPlot(sce, vars = markers , group.by = "label",min.color="lightgrey", max.color="blue",min.percent=0.2,max.percent=1,scale = TRUE, y.reorder=c(4,9,3,6,7,2,5,1,8,10))

#order-->c("Fibroblast","Smooth muscle cell","Endothelial cell","Myeloid","Neutrophil","Dendritic cell","Mast cell","B cell","Plasma","T cell")

rm(sce.plot,column_annotations,markers_wilcox_up)
gc()
```


## Frequencies of celltypes in clusters per Sample {.tabset}

```{r ClusterInfoSample}
ClusterInfo <- as.data.frame(colData(sce)) %>% 
  group_by(celltype, Sample,Joint.Location) %>%
  summarise(cells = n(),.groups = 'drop') 
```

### Absolute comparison

```{r PlotAbsoluteClusteringSample}

ggplot(data=ClusterInfo, aes(x=Sample,y=cells, fill=celltype)) +
  geom_bar(stat="identity", position="stack") +
  scale_fill_manual( values=meta_colors$celltype_level1[seq.int(length(unique(ClusterInfo$celltype)))] )+
  labs(title="Absolute comparison nr Cells in Celltype",subtitle="Per Sample",x="",y="Number of cells")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())+
  scale_y_continuous(expand = c(0,0)) 

```

### Relative comparison

```{r PlotRelativeClusteringSample}

ggplot(data=ClusterInfo,aes(x=Sample, y=cells, fill=celltype)) +
  geom_bar(stat="identity", position="fill") +
  scale_fill_manual( values=meta_colors$celltype_level1[seq.int(length(unique(ClusterInfo$celltype)))] )+
  labs(title="Relative comparison nr Cells in celltype",subtitle="Per Sample",x="",y="Number of cells")+
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())+
  scale_y_continuous(expand = c(0,0)) 

```

## Frequencies of celltypes in clusters per Condition {.tabset}

```{r ClusterInfoCond}
ClusterInfo <- as.data.frame(colData(sce)) %>% 
  group_by(celltype,Joint.Location) %>%
  summarise(cells = n(),.groups = 'drop') 
```

### Absolute comparison

```{r PlotAbsoluteClusteringCond}

ggplot(data=ClusterInfo, aes(x=Joint.Location,y=cells, fill=celltype)) +
  geom_bar(stat="identity", position="stack") +
  scale_fill_manual( values=meta_colors$celltype_level1[seq.int(length(unique(ClusterInfo$celltype)))] )+
  labs(title="Absolute comparison nr Cells in Celltype",subtitle="Per Condition",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())+
  theme_classic()

```

### Relative comparison

```{r PlotRelativeClusteringCond}

ggplot(data=ClusterInfo,aes(x=Joint.Location, y=cells, fill=celltype)) +
  geom_bar(stat="identity", position="fill") +
  scale_fill_manual( values=meta_colors$celltype_level1[seq.int(length(unique(ClusterInfo$celltype)))] )+
  labs(title="Relative comparison nr Cells in celltype",subtitle="Per Condition",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())+
  theme_classic()


```


### Grouped boxplot

```{r PlotGroupedClustering}
# Fraction of cells [%] from total cells in that celltype
Totaljoint <- as.data.frame(colData(sce)) %>% 
     group_by(Joint.Location) %>%
     summarise(totalcellsjoint = n(),.groups = 'drop')

Totalcelltype <- as.data.frame(colData(sce)) %>% 
    group_by(Joint.Location,celltype) %>%
    summarise(totalcells = n(),.groups = 'drop') %>% 
    left_join(Totaljoint, Totalcelltype,by ="Joint.Location") %>%
    mutate(freq = totalcells / totalcellsjoint * 100)


ggplot(Totalcelltype, aes(fill=Joint.Location, y=freq, x=celltype)) + 
    geom_bar(width=0.8,position=position_dodge(0.9), stat="identity")+ 
    scale_fill_manual( values=meta_colors$Joint.Location[seq.int(length(unique(ClusterInfo$celltype)))] )+
    theme_classic()+
    labs(title="Relative cell type abundance in celltypes", subtitle="Per Joint Location", x="", y="Fraction of cells [%]")+
    theme(axis.text.x = element_text(angle = 45,hjust=1), axis.line = element_line(colour = "black"), panel.background = element_rect(fill = NA)) 



```




## Identify Marker Genes for Annotation


### Wilcoxon rank-sum test on Annotation

```{r WilcoxonLabel}
for (labeling in c('Annotation_L0','Annotation_L1')){
  #Wilcoxon rank-sum test
  markers_wilcox <- wilcoxauc(sce, group_by=labeling, assay = 'logcounts')
  topDEgenes <- top_markers(markers_wilcox, n = 100, auc_min = 0.5, pval_max = 0.05)# filter for significant (p<0.05) and over-expressed (auc>0.5)
  wb = createWorkbook()
  for (i in unique(markers_wilcox$group) ){
    #get genes for the Anno
    markers<-topDEgenes[,paste0(i), drop=TRUE]
    #Create table
    tmp<-markers_wilcox[markers_wilcox$feature %in% markers & markers_wilcox$group == i,]
    #Save table in Exelsheet
    addDataFrame(tmp, sheet=createSheet(wb, paste0(i)), startColumn=1, row.names=FALSE,col.names = TRUE)
  }
  saveWorkbook(wb,paste0( output_dir , "/",paste0(labeling),"_Wilcoxon.xlsx"))  
}

```


### Assignement

```{r AssignementL}
## Set Celltypes to use downstream as colLabels
colLabels(sce) <- colData(sce)$Annotation_L1
sce$celltype <- colData(sce)$Annotation_L1
```


## Save the dataset

```{r SaveData}
saveRDS(sce, file =paste0(path,'/output/06_sce_CelltypeAnnotation.rds'))
```
