---
title: "Subcluster Fibroblast"
subtitle: "06_SC_Fibroblast"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Subcluster Fibroblast

## Setup

### Standard packages

```{r LoadPackages, warning=FALSE}
library(here)
source(here("code", "standard_libraries.R"))
```

### Additional Packages


```{r LoadPackages, warning=FALSE,eval=FALSE}
#Load Packages

suppressPackageStartupMessages({
library(batchelor)
library(bluster)
library(presto)
library(dittoSeq)
library(gridExtra)
})

```

### Set Parameter

```{r Setup}
set.seed(100)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=100)
path <- here::here()
```


### Load Data

```{r LoadDataset}
## RA DATASET
sce <- readRDS(file = paste0(path,'/output/06_sce_SC_Fibroblast_Anno.rds'))

```

### Explore Dataset

```{r ShowDim}
#Dimensions of count matrix
dim(sce)
```

https://mperalc.gitlab.io/scRNA-seq_workshop_2021/additional-resources.html## UMAPs coloured by cluster or cell type

### Plot Annotation {.tabset}

#### Plot UMAP of Subclusters 

```{r UMAPCluster}
#cluster
plotReducedDim(sce, paste0("MNN_UMAP_SC_reduced"),colour_by = "subcluster", text_by="subcluster") +labs(title=paste0("SubCluster"),subtitle="Colored by clustering")
```

#### Plot Recalculated UMAP of Subclusters 

```{r UMAPCluster}
#cluster
plotReducedDim(sce, paste0("MNN_UMAP_SC_reduced"),colour_by = "subcluster", text_by="subcluster") +labs(title="SubCluster ",subtitle="Colored by clustering")
```


## DA

### Frequencies of celltypes in clusters per Sample {.tabset}

```{r ClusterInfo}
ClusterInfo <- as.data.frame(colData(sce)) %>% 
  group_by(ManualAnno_L2, Sample) %>%
  summarise(cells = n(),.groups = 'drop') 
```

#### Absolute comparison

```{r PlotAbsoluteClustering}

ggplot(data=ClusterInfo, aes(x=Sample,y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="stack") +
  labs(title="Absolute comparison nr Cells in Celltype",subtitle="Per Sample",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())

```

#### Relative comparison

```{r PlotRelativeClustering}

ggplot(data=ClusterInfo,aes(x=Sample, y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="fill") +
  labs(title="Relative comparison nr Cells in celltype",subtitle="Per Sample",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())


```

### Frequencies of celltypes in clusters per Joint Location {.tabset}

```{r ClusterInfo}
ClusterInfo <- as.data.frame(colData(sce)) %>% 
  group_by(ManualAnno_L2, Joint.Location) %>%
  summarise(cells = n(),.groups = 'drop') 
```

#### Absolute comparison

```{r PlotAbsoluteClustering}

ggplot(data=ClusterInfo, aes(x=Joint.Location,y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="stack") +
  labs(title="Absolute comparison nr Cells in Celltype",subtitle="Per Joint Location",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())

```

#### Relative comparison

```{r PlotRelativeClustering}

ggplot(data=ClusterInfo,aes(x=Joint.Location, y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="fill") +
  labs(title="Relative comparison nr Cells in celltype",subtitle="Per Joint Location",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())


```

```{r TestProportionsScProportionTest}
library(scProportionTest)
library(Seurat)

so <- as.Seurat(sce)

prop_test <- sc_utils(so)

#Knee vs MCP
prop_test <- permutation_test(
	prop_test, cluster_identity = "ManualAnno_L2",
	sample_1 = "Knee", sample_2 = "MCP",
	sample_identity = "Joint.Location"
)
print(prop_test)
permutation_plot(prop_test)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())


#Knee vs Wrist

prop_test <- permutation_test(
	prop_test, cluster_identity = "ManualAnno_L2",
	sample_1 = "Knee", sample_2 = "Wrist",
	sample_identity = "Joint.Location"
)

print(prop_test)
permutation_plot(prop_test)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())


#MCP vs Wrist

prop_test <- permutation_test(
	prop_test, cluster_identity = "ManualAnno_L2",
	sample_1 = "MCP", sample_2 = "Wrist",
	sample_identity = "Joint.Location"
)
print(prop_test)
permutation_plot(prop_test)+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())



```




```{r }


Marker <- c("HOXD11","HOXD10","HOXB-AS1","HOXA6","HOXA5","HOXA3","HOXA11-AS","HOXC9","HOXC8","HOXC11","HOXC10","HOXC-AS1","HOXC-AS2","HOXC-AS3","HOXB3","HOXB2")


dittoDotPlot(sce[,sce$celltype == "Fibroblast"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = FALSE )

dittoDotPlot(sce[,sce$celltype == "Fibroblast"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = TRUE )



dittoDotPlot(sce[,sce$celltype == "Edothelial cell"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = FALSE )

dittoDotPlot(sce[,sce$celltype == "Edothelial cell"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = TRUE )



```

## DE

-MAST between the clusters


## Trajectory analysis

### Plot UMAP of Subclusters 

```{r UMAPCluster}
#cluster
plotReducedDim(sce, paste0("MNN_UMAP_SC_reduced"),colour_by = "subcluster", text_by="subcluster") +labs(title=paste0("SubCluster"),subtitle="Colored by clustering")
plotReducedDim(sce, paste0("MNN_SC_reduced"),colour_by = "subcluster", text_by="subcluster") +labs(title=paste0("SubCluster"),subtitle="Colored by clustering")
```


```{r }
library(slingshot)
library(ggbeeswarm)
library(tradeSeq)
#sce<-sce[,sce$subcluster != 6]


```


# Trajectory Inference 

```{r }

sce <- slingshot(sce, reducedDim = 'MNN_SC_reduced')


slingshot_df <- data.frame(colData(sce)[, names(colData(sce)) != 'slingshot', drop=FALSE])

ggplot(slingshot_df, aes(x = slingPseudotime_1, y = subcluster, 
                              colour = subcluster)) +
    geom_quasirandom(groupOnX = FALSE) + theme_classic() +
    xlab("First Slingshot pseudotime") + ylab("cell type") +
    ggtitle("Cells ordered by Slingshot pseudotime")


```


# Differential progression
```{r plotSlingshot, echo = FALSE, fig.width =  8}

shuffle <- sample(ncol(sce))
layout(matrix(1:2, nrow = 1))
par(mar = c(4.5,4,1,1))



layout(matrix(c(1, 1, 2, 3), 2))
par(mar = c(4.5, 4, 1, 1))
plot(reducedDims(sce)$MNN_SC_reduced[shuffle, ], asp = 1, pch = 16, xlab = "UMAP-1", ylab = "UMAP-2",
  col = hcl.colors(100, alpha = .5)[cut(sce$slingPseudotime_1, breaks = 100)][shuffle])
lines(SlingshotDataSet(sce))


library(RColorBrewer)
# Pseudotime densities (by Joint) -- Differential progression
ds <- list( MCP = density(slingPseudotime(sce)[colData(sce)$Joint.Location == "MCP", 1]),
            Wrist = density(slingPseudotime(sce)[colData(sce)$Joint.Location == "Wrist", 1]),
            Knee = density(slingPseudotime(sce)[colData(sce)$Joint.Location == "Knee", 1])
            )
xlim <- range(c(ds$MCP$x, ds$Wrist$x, ds$Knee$x))
ylim <- range(c(ds$MCP$y, ds$Wrist$y, ds$Knee$y))
plot(xlim, ylim, col = "white", xlab = "Pseudotime", ylab = "")
polygon(c(min(ds$MCP$x), ds$MCP$x, max(ds$MCP$x)), c(0, ds$MCP$y, 0),
  col = alpha(brewer.pal(4, "Set1")[3], alpha = .5))
polygon(c(min(ds$Wrist$x), ds$MCP$x, max(ds$MCP$x)), c(0, ds$Wrist$y, 0),
        col = alpha(brewer.pal(4, "Set1")[2], alpha = .5))
polygon(c(min(ds$Knee$x), ds$MCP$x, max(ds$MCP$x)), c(0, ds$Knee$y, 0),
  col = alpha(brewer.pal(4, "Set1")[1], alpha = .5))

legend("topleft", legend = c("MCP", "Wrist", "Knee"), 
       fill = alpha(brewer.pal(4, "Set1")[3:1], alpha = .5), bty = "n")

plot(reducedDims(sce)$MNN_SC_reduced[shuffle, ], asp = 1, pch = 16, xlab = "UMAP-1", ylab = "UMAP-2", 
     col = alpha(c(2, 3, 4)[factor(colData(sce)$Joint.Location)][shuffle], alpha = .5))
lines(SlingshotDataSet(sce), type = 'lineages', show.constraints = TRUE)
legend("topright", pch = 16, col = 2:4, bty = "n", legend = levels(factor(colData(sce)$Joint.Location)))

layout(1)
par(mar = c(5, 4, 4, 2) + .1)



```

```{r plotSlingshot, echo = FALSE, fig.width =  8}
ks.test(slingPseudotime(sce)[colData(sce)$Joint.Location == "Knee", 1],
        slingPseudotime(sce)[colData(sce)$Joint.Location == "MCP", 1])

ks.test(slingPseudotime(sce)[colData(sce)$Joint.Location == "Knee", 1],
        slingPseudotime(sce)[colData(sce)$Joint.Location == "Wrist", 1])

ks.test(slingPseudotime(sce)[colData(sce)$Joint.Location == "MCP", 1],
        slingPseudotime(sce)[colData(sce)$Joint.Location == "Wrist", 1])
```


# Differential expression
```{r }
plotExpression(sce, features="HOXD10", x = "slingPseudotime_1", show_violin = FALSE, show_smooth = TRUE, colour_by = "subcluster" )
```


## Select number of knots

```{r }
set.seed(123)
sdset <- SlingshotDataSet(sce)

icMat <- evaluateK(counts = as.matrix(assays(sce)$counts),
                   pseudotime = colData(sce)$slingshot$pseudotime,
                   sds=sdset,
                   conditions = factor(colData(sce)$Joint.Location),
                   nGenes = 300,
                   k = 3:7,
                   BPPARAM=bpp)

```

# Fit GAM
```{r }

set.seed(123)
sce <- fitGAM(sce, conditions = factor(colData(sce)$Joint.Location),
              nknots = 5, BPPARAM = BPPARAM)
mean(rowData(sce)$tradeSeq$converged)

```





--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------


## Save the dataset

```{r SaveData}
saveRDS(sce, file =paste0(path,'/output/06_sce_SC_Fibroblast.rds'))
```
