---
title: "Subcluster Fibroblast"
subtitle: "06_SC_Fibroblast"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Subcluster Fibroblast

## Setup

### Standard packages

```{r LoadPackages, warning=FALSE}
library(here)
source(here("code", "standard_libraries.R"))
```

### Additional Packages


```{r LoadPackages, warning=FALSE,eval=FALSE}
#Load Packages

suppressPackageStartupMessages({
library(batchelor)
library(bluster)
library(presto)
library(dittoSeq)
library(gridExtra)
})

```

```{r LoadPackages, warning=FALSE}
#Load Packages
library(here)
source(here("code", "standard_libraries.R"))

```

### Set Parameter

```{r Setup}
set.seed(100)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=100)
path <- here::here()
```


### Load Data

```{r LoadDataset}
## RA DATASET
sce <- readRDS(file = paste0(path,'/output/05_sce_CelltypeAnnotation.rds'))

```

### Explore Dataset

```{r ShowDim}
#Dimensions of count matrix
dim(sce)
```

https://mperalc.gitlab.io/scRNA-seq_workshop_2021/additional-resources.html## UMAPs coloured by cluster or cell type

### Plot Annotation {.tabset}

#### Plot UMAP of Subclusters 

```{r UMAPCluster}
#cluster
plotReducedDim(sce, paste0("MNN_UMAP_reduced", cell_type),colour_by = "subcluster", text_by="subcluster") +labs(title=paste0("SubCluster ", name),subtitle="Colored by clustering")
```

#### Plot Recalculated UMAP of Subclusters 

```{r UMAPCluster}
#cluster
plotReducedDim(sce, paste0("SC_UMAP_", cell_type),colour_by = "subcluster", text_by="subcluster") +labs(title=paste0("SubCluster ", name),subtitle="Colored by clustering")
```



## Annotate Subcluster


## Annotate Subclusters 

```{r SetupAssignement}
#Manual Annotation setup
sce.sc$ManualAnno_L2 <- as.integer(sce.sc$celltype)
```

```{r VizGenExFunction}
#Function to visuaise gene Expression for vector of genes

VizGenEx <- function(markr,Ctype){
  for(i in markr){
    
    cat(paste0("#### ", i, "\n"))
    gridExtra::grid.arrange(
      arrangeGrob(plotReducedDim(sce.sc[,sce.sc$celltype == Ctype], dimred = paste0("SC_UMAP_",Ctype), colour_by = i, text_by = "subcluster"), 
                  plotReducedDim(sce.sc[,sce.sc$celltype == Ctype], dimred = paste0("SC_UMAP_",Ctype), colour_by = i, order_by=i ) 
                    +scale_colour_gradient(name = i,low = "lightgrey", high = "red"), ncol = 2), 
                  plotExpression(sce.sc[,sce.sc$celltype == Ctype], features = i, x = "subcluster"),nrow = 2)
    cat(' \n')
  
  }
}

```


#### Identify Marker Genes Wilcoxon rank-sum test

```{r FibroblastsIdentifyMarker}
markers_wilcox <- wilcoxauc(sce.sc[,sce.sc$celltype == "Fibroblast"], 'subcluster', assay = 'logcounts')
data.frame(markers_wilcox)
```

```{r FibroblastsPlotMarker}
tmp <-top_markers(markers_wilcox, n = 4)

tail(unlist(unname(as.list(tmp))),-4)

dittoDotPlot(sce.sc[,sce.sc$celltype == "Fibroblast"], vars = unique(tail(unlist(unname(as.list(tmp))),-4)) , group.by = "subcluster",min.color="lightgrey", max.color="blue",min.percent=0.2,max.percent=1,scale = TRUE )
#dittoDotPlot(sce.sc[,sce.sc$celltype == "Fibroblast"], vars =unique(tail(unlist(unname(as.list(tmp))),-4)) , group.by = "subcluster", min.color="white", max.color="red")
```


#### Test Fibroblast marker {.tabset}

Lining vs Sublining


```{r FibroblastsL2, results='asis'}
#Lining Fibroblst
VizGenEx(c("PRG4","THY1"),"Fibroblast")  


```


```{r FibroblastsL2, results='asis'}
#Lining Fibroblst
VizGenEx(c("PRG4","CRTAC1", "CLU","CLIC5"),"Fibroblast")  


```



```{r FibroblastsL2, results='asis'}
#SMC
VizGenEx(c("ACTA2","MYL9", "DES", "MYH11"),"Fibroblast")  


```


```{r FibroblastsL2, results='asis'}
#Lining Fibroblst
VizGenEx(c("CHI3L2","SAA1", "FN1"),"Fibroblast")  


```

```{r FibroblastsL2, results='asis'}
#Lining Fibroblst
VizGenEx(c("CXCL12","CCL2","SFRP1"),"Fibroblast")  


```


```{r FibroblastsL2, results='asis'}
#Lining Fibroblst
VizGenEx(c("POSTN","COL1A1", "FNDC1"),"Fibroblast")  


```


```{r FibroblastsL2, results='asis'}
#Lining Fibroblst
VizGenEx(c("CXCL14","MFAP5"),"Fibroblast")  


```

```{r FibroblastsL2, results='asis'}
#Lining Fibroblst
VizGenEx(c("TRPS1","FTX","NEAT1","MALAT1"),"Fibroblast")  

```


```{r FibroblastsAssignementL2}
#Manual Annotation


sce.sc$ManualAnno_L2 <- as.integer(sce.sc$subcluster)
unique(sce$ManualAnno_L2)

sce.sc$ManualAnno_L2[sce.sc$subcluster %in% c(6)] <- "SMC"
sce.sc$ManualAnno_L2[sce.sc$subcluster %in% c(1,8)] <- "POSTN+"
sce.sc$ManualAnno_L2[sce.sc$subcluster %in% c(2,7)] <- "CHI3L2+"
sce.sc$ManualAnno_L2[sce.sc$subcluster %in% c(5)] <- "PRG4+"
sce.sc$ManualAnno_L2[sce.sc$subcluster %in% c(3)] <- "CXCL12+"
sce.sc$ManualAnno_L2[sce.sc$subcluster %in% c(4)] <- "FTX+"

sce$ManualAnno_L2 <- as.integer(sce.sc$ManualAnno_L2)
```

#### UMAP Assigned Celltypes 

```{r FibroblastsUMAPAssigned}
plotReducedDim(sce.sc[,sce.sc$celltype == name], column_name,colour_by = "ManualAnno_L2",text_by="ManualAnno_L2") +
          labs(title=paste0("Recalculated UMAP for Subcluster ", name),subtitle="Colored by Celltype")
```

```{r FibroblastsUMAPAssigned}
plotReducedDim(sce.sc[,sce.sc$celltype == name], column_name,colour_by = "ManualAnno_L2") +
          labs(title=paste0("Recalculated UMAP for Subcluster ", name),subtitle="Colored by Celltype")
```



## Frequencies of celltypes in clusters per Sample {.tabset}

```{r ClusterInfo}
ClusterInfo <- as.data.frame(colData(sce.sc[,sce.sc$celltype == "Fibroblast"])) %>% 
  group_by(ManualAnno_L2, Sample) %>%
  summarise(cells = n(),.groups = 'drop') 
```

### Absolute comparison

```{r PlotAbsoluteClustering}

ggplot(data=ClusterInfo, aes(x=Sample,y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="stack") +
  labs(title="Absolute comparison nr Cells in Celltype",subtitle="Per Sample",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())

```

### Relative comparison

```{r PlotRelativeClustering}

ggplot(data=ClusterInfo,aes(x=Sample, y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="fill") +
  labs(title="Relative comparison nr Cells in celltype",subtitle="Per Sample",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())


```


## Frequencies of celltypes in clusters per Sample {.tabset}

```{r ClusterInfo}
ClusterInfo <- as.data.frame(colData(sce.sc[,sce.sc$celltype == "Fibroblast"])) %>% 
  group_by(ManualAnno_L2, Joint.Location) %>%
  summarise(cells = n(),.groups = 'drop') 
```

### Absolute comparison

```{r PlotAbsoluteClustering}

ggplot(data=ClusterInfo, aes(x=Joint.Location,y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="stack") +
  labs(title="Absolute comparison nr Cells in Celltype",subtitle="Per Joint Location",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())

```

### Relative comparison

```{r PlotRelativeClustering}

ggplot(data=ClusterInfo,aes(x=Joint.Location, y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="fill") +
  labs(title="Relative comparison nr Cells in celltype",subtitle="Per Joint Location",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())


```

```{r TestProportionsWilcoxon}

ClusterInfo
wilcox.test()




```


## DA


```{r }






Marker <- c("HOXD11","HOXD10","HOXB-AS1","HOXA6","HOXA5","HOXA3","HOXA11-AS","HOXC9","HOXC8","HOXC11","HOXC10","HOXC-AS1","HOXC-AS2","HOXC-AS3","HOXB3","HOXB2")


dittoDotPlot(sce[,sce$celltype == "Fibroblast"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = FALSE )

dittoDotPlot(sce[,sce$celltype == "Fibroblast"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = TRUE )



dittoDotPlot(sce[,sce$celltype == "Edothelial cell"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = FALSE )

dittoDotPlot(sce[,sce$celltype == "Edothelial cell"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = TRUE )



```

## DE





--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------


## Save the dataset

```{r SaveData}
saveRDS(sce, file =paste0(path,'/output/06_sce_SC_Fibroblast.rds'))
```
