---
title: "Subclustering"
subtitle: "06_SubClustering"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Subclustering

## Setup

### Standard packages

```{r LoadPackages, warning=FALSE}
library(here)
source(here("code", "standard_libraries.R"))
```

### Additional Packages


```{r LoadPackages, warning=FALSE,eval=FALSE}
#Load Packages

suppressPackageStartupMessages({
library(batchelor)
library(bluster)
library(presto)
library(dittoSeq)
library(gridExtra)
})

```

```{r LoadPackages, warning=FALSE}
#Load Packages
library(here)
source(here("code", "standard_libraries.R"))

```

### Set Parameter

```{r Setup}
set.seed(100)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=100)
path <- here::here()
```


### Load Data

```{r LoadDataset}
## RA DATASET
sce <- readRDS(file = paste0(path,'/output/05_sce_CelltypeAnnotation.rds'))

```


### Explore Dataset

```{r ShowDim}
#Dimensions of count matrix
dim(sce)
```

https://mperalc.gitlab.io/scRNA-seq_workshop_2021/additional-resources.html## UMAPs coloured by cluster or cell type

### Plot Annotation {.tabset}

### Annotation L1

```{r UMAPCluster}
#cluster
plotReducedDim(sce, "MNN_UMAP", colour_by="ManualAnno_L1", text_by="ManualAnno_L1") +labs(title="UMAP colored by clusters",subtitle = "UMAP of integrated dataset")

```

### Annotation L0

```{r UMAPAnnotation}
#cell type
plotReducedDim(sce, "MNN_UMAP", colour_by="ManualAnno_L0", text_by="ManualAnno_L0") +labs(title="UMAP colored by Cell type ",subtitle = "UMAP of integrated dataset")

```

## Subclustering

```{r SubclusteringFunction}
# Create Subclustering function

SUBCLUSTER <- function(sce.sc, SC_celltypes, clust_param="leiden", k_param=50){

  for ( name in SC_celltypes){
  
    #Subset sce object for ce scelltype
    sce.subset <- sce.sc[,sce.sc$celltype == name]
    
    #Get top HVG for subcluster
    dec <- modelGeneVar(sce.subset, block=sce.subset$Sample, density.weights=FALSE,BPPARAM=bpp)
    top.hvgs <- getTopHVGs(dec, n=3000)
    rowData(sce.subset)$is_hvg <- rownames(sce.subset) %in% top.hvgs
  
    
    #Batch correction on subcluster
    sce.Batchelor <-multiBatchNorm(sce.subset, 
                                   batch=sce.subset$Sample, 
                                   subset.row = rownames(sce.subset)[rowData(sce.subset)[["is_hvg"]]], 
                                   normalize.all=TRUE, BPPARAM = bpp)
    
    bpstart(bpp)
    sce.Batchelor <- fastMNN(sce.Batchelor, 
                             batch=sce.Batchelor$Sample, 
                             prop.k=0.05,  
                             subset.row = rownames(sce.Batchelor)[rowData(sce.Batchelor)[["is_hvg"]]], 
                             correct.all=TRUE, BPPARAM = bpp)#merge.order = merge_order,
    bpstop(bpp)
    
    assay(sce.subset, "reconstructed") <- assay(sce.Batchelor, "reconstructed")
    reducedDim(sce.subset, 'MNN') <- reducedDim(sce.Batchelor, 'corrected')
    
    #Recluster the cells 
    clusters <- clusterCells(sce.subset, use.dimred = "MNN", BLUSPARAM = SNNGraphParam(k= k_param,cluster.fun= clust_param, BPPARAM =bpp))
    
    #Add Subcluster label to sce
    colData(sce.sc)[sce.sc$celltype == name, "subcluster"] <- clusters
  
    #Get UMAP on subset
    sce.subset <- runUMAP(sce.subset, dimred = 'MNN', subset_row=rowData(sce.subset)$is_hvg, min_dist = 0.1)
    umap_for_cell_type <- reducedDim(sce.subset, "UMAP")
    column_name <- paste0("SC_UMAP_", name)
  
    # Add dimension reduction for only selected subcluster cells
    reducedDim(sce.sc, column_name) <- reducedDim(sce.sc, "MNN_UMAP")
    reducedDim(sce.sc, column_name)[,c(1,2)] <- NA
  
    attr(reducedDim(sce.sc, column_name), "scaled:center") <- attr(umap_for_cell_type, "scaled:center")
    reducedDim(sce.sc, column_name)[rownames(umap_for_cell_type),] <- umap_for_cell_type
  }
  
  # transform to factor values
  sce.sc$subcluster_complete <- factor(paste0(sce.sc$celltype, "_", sce.sc$subcluster))
  sce.sc$subcluster <- factor(sce.sc$subcluster)
  
  # Add column for each of the cell type in ColData
  for (name in SC_celltypes){
    colData(sce.sc)[,name] <- colData(sce.sc)[,"subcluster"]
    colData(sce.sc)[sce.sc$celltype != name, name] <- NA
    colData(sce.sc)[, name] <- factor(as.numeric(colData(sce.sc)[, name]))
  }
  
  return(sce.sc)
}
```

```{r RunSubcluster, warning=FALSE}

#Set Celltypes to be subclustered
SC_celltypes <-c("Fibroblast","Tcell")
#SC_celltypes <-unique(sce$celltype) #All celltypes

sce$celltype <- colData(sce)$ManualAnno_L1

# Run Subcluster Function
sce.sc<-SUBCLUSTER(sce, SC_celltypes, clust_param="leiden", k_param=40)
  
```

## Plot Subclusters

### Plot UMAP of Subclusters  {.tabset}

```{r UMAPSubcluster}
# plot UMAP of the subclusters

for (name in SC_celltypes){
  cat(paste0("#### ", SC_celltypes, " {.tabset} \n"))
  
  cat(paste0("##### Colored by Clustering\n"))
  print(plotReducedDim(sce.sc[,sce.sc$celltype == name], "MNN_UMAP",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("SubCluster ", name),subtitle="Colored by clustering"))
    
  cat(paste0("##### Colored by Sample\n"))
  print(plotReducedDim(sce.sc[,sce.sc$celltype == name], "MNN_UMAP",colour_by = "Sample") +
          labs(title=paste0("SubCluster ", name),subtitle="Colored by Sample"))
}

```

### Plot Recalculated UMAP of Subclusters  {.tabset}

```{r UMAPSubclusterRecalc}
# plot the Recalculated UMAP of the subcluster

for (name in SC_celltypes){
  cat(paste0("#### ", SC_celltypes, " {.tabset} \n"))  
  
  column_name <- paste0("SC_UMAP_", name)
  
  cat(paste0("##### Colored by Clustering\n"))
  print(plotReducedDim(sce.sc[,sce.sc$celltype == name], column_name,colour_by = "subcluster",text_by="subcluster") +
          labs(title=paste0("Recalculated UMAP for Subcluster ", name),subtitle="Colored by clustering"))

  cat(paste0("##### Colored by Sample\n"))
  print(plotReducedDim(sce.sc[,sce.sc$celltype == name], column_name,colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster ", name),subtitle="Colored by Sample"))
}

```


## Annotate Subclusters 

```{r SetupAssignement}
#Manual Annotation setup
sce.sc$ManualAnno_L2 <- as.integer(sce.sc$celltype)
```

```{r VizGenExFunction}
#Function to visuaise gene Expression for vector of genes

VizGenEx <- function(markr,Ctype){
  for(i in markr){
    
    cat(paste0("#### ", i, "\n"))
    gridExtra::grid.arrange(
      arrangeGrob(plotReducedDim(sce.sc[,sce.sc$celltype == Ctype], dimred = paste0("SC_UMAP_",Ctype), colour_by = i, text_by = "subcluster"), 
                  plotReducedDim(sce.sc[,sce.sc$celltype == Ctype], dimred = paste0("SC_UMAP_",Ctype), colour_by = i, order_by=i ) 
                    +scale_colour_gradient(name = i,low = "lightgrey", high = "red"), ncol = 2), 
                  plotExpression(sce.sc[,sce.sc$celltype == Ctype], features = i, x = "subcluster"),nrow = 2)
    cat(' \n')
  
  }
}

```

--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

### Fibroblasts

```{r FibroblastsUMAP}
#Plot the UMAp
plotReducedDim(sce.sc[,sce.sc$celltype == "Fibroblast"], "SC_UMAP_Fibroblast" ,colour_by = "subcluster",text_by="subcluster") +
          labs(title=paste0("Recalculated UMAP for Subcluster ", "Fibroblast"),subtitle="Colored by clustering")
```

#### Identify Marker Genes Wilcoxon rank-sum test

```{r FibroblastsIdentifyMarker}
markers_wilcox <- wilcoxauc(sce.sc[,sce.sc$celltype == "Fibroblast"], 'subcluster', assay = 'logcounts')
data.frame(markers_wilcox)
```

```{r FibroblastsPlotMarker}
tmp <-top_markers(markers_wilcox, n = 4)

tail(unlist(unname(as.list(tmp))),-4)

dittoDotPlot(sce.sc[,sce.sc$celltype == "Fibroblast"], vars =unique(tail(unlist(unname(as.list(tmp))),-4)) , group.by = "subcluster", min.color="white", max.color="red")
```


#### Test Fibroblast marker {.tabset}

```{r FibroblastsL2, results='asis'}
#Lining Fibroblst
VizGenEx(c("PRG4","CRTAC1", "CLU"),"Fibroblast")  


```


```{r FibroblastsAssignementL2}
#Manual Annotation

sce$ManualAnno_L2 <- as.integer(sce$subcluster)
unique(sce$ManualAnno_L2)

sce$ManualAnno_L2[sce$subcluster %in% c(2,3,12)] <- "Fibroblast"

```

#### UMAP Assigned Celltypes 

```{r FibroblastsUMAPAssigned}
plotReducedDim(sce.sc[,sce.sc$celltype == name], column_name,colour_by = "ManualAnno_L2",text_by="ManualAnno_L2") +
          labs(title=paste0("Recalculated UMAP for Subcluster ", name),subtitle="Colored by Celltype")
```

--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

### Tcells


```{r TcellUMAP}
#Plot the UMAp
plotReducedDim(sce.sc[,sce.sc$celltype == "Tcell"], "SC_UMAP_Tcell" ,colour_by = "subcluster",text_by="subcluster") +
          labs(title=paste0("Recalculated UMAP for Subcluster ", "Tcell"),subtitle="Colored by clustering")
```

#### Identify Marker Genes Wilcoxon rank-sum test

```{r TcellIdentifyMarker}
markers_wilcox <- wilcoxauc(sce.sc[,sce.sc$celltype == "Tcell"], 'subcluster', assay = 'logcounts')
data.frame(markers_wilcox)
```

```{r TcellPlotMarker}
tmp <-top_markers(markers_wilcox, n = 4)

tail(unlist(unname(as.list(tmp))),-4)

dittoDotPlot(sce.sc[,sce.sc$celltype == "Fibroblast"], vars =unique(tail(unlist(unname(as.list(tmp))),-4)) , group.by = "subcluster", min.color="white", max.color="red")
```

#### Test Tcells marker {.tabset}

```{r TcellL2, results='asis'}
#CD8Tcell
VizGenEx(c("CD4","TRAT1", "CD40LG", "TNFRSF4"),"Tcell")  


```

```{r TcellL2, results='asis'}
#CD4Tcell
VizGenEx(c("CD8A","CD8B"),"Tcell")  


```

```{r TcellL2, results='asis'}
#NKcell
VizGenEx(c("NKG7","GNLY","NCAM1"),"Tcell")  


```
```{r TcellL2, results='asis'}
#ProlifTcell
VizGenEx(c("MKI67","PCNA","STMN1"),"Tcell")  


```

```{r TcellAssignementL2}
#Manual Annotation

sce$ManualAnno_L2 <- as.integer(sce$subcluster)
unique(sce$ManualAnno_L2)

sce$ManualAnno_L2[sce$subcluster %in% c(4)] <- "NKcell"
sce$ManualAnno_L2[sce$subcluster %in% c(6)] <- "CD4Tcell"
sce$ManualAnno_L2[sce$subcluster %in% c(5)] <- "CD8Tcell"
sce$ManualAnno_L2[sce$subcluster %in% c(1)] <- "ProlifTcell"
```

#### UMAP Assigned Celltypes 

```{r TcellUMAPAssigned}
plotReducedDim(sce.sc[,sce.sc$celltype == name], column_name,colour_by = "ManualAnno_L2",text_by="ManualAnno_L2") +
          labs(title=paste0("Recalculated UMAP for Subcluster ", name),subtitle="Colored by Celltype")
```


--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

### Myeloid


```{r TcellL2, results='asis'}
#Macrophage
VizGenEx(c("CD14","IL1B","LYZ","CD163","ITGAX","CD68","CSF1R","FCGR3A"),"Myeloid")  


```


```{r TcellL2, results='asis'}
#M1 Macrophage
VizGenEx(c("CD38","Gpr18","Fpr2"),"Myeloid")  


```


```{r TcellL2, results='asis'}
#M2 Macrophage
VizGenEx(c("Egr2","C-MYC"),"Myeloid")  


```

```{r TcellL2, results='asis'}
#monocyte
VizGenEx(c("CST3","CSF1R","ITGAM","CD14","FCGR3A","FCGR3B"),"Myeloid")  


```

```{r TcellL2, results='asis'}
#cDC
VizGenEx(c("FLT3","UPK3A","CLEC9A","CCL22","ZNF366","CD1C"),"Myeloid")  

```

--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

### Endothelial


```{r TcellL2, results='asis'}
#Capillary
VizGenEx(c("SPARC"),"Endothelial")  

```


```{r TcellL2, results='asis'}
#Venular
VizGenEx(c("LIFR","ICAM"),"Endothelial")  

```


```{r TcellL2, results='asis'}
#Arteriolar
VizGenEx(c("NOTCH","GAJ5","B2MX"),"Endothelial")  

```


```{r TcellL2, results='asis'}
#lymphatic
VizGenEx(c("PROX1","LYVE1", "CCL21"),"Endothelial")  

```


--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------



```{r AssignementCheck}
#Check if all clusters assigned
unique(sce$ManualAnno_L2)
```


### Plot Assigned Celltypes {.tabset}

##### UMAP All Subcluster

```{r PlotL0}
#Plot manual Annotation
plotReducedDim(sce, dimred="MNN_UMAP", by_exprs_values = "logcounts", colour_by="ManualAnno_L2")+ ggtitle("UMAP Celllabel")
plotReducedDim(sce, dimred="MNN_UMAP", by_exprs_values = "logcounts", colour_by="ManualAnno_L2", text_by="ManualAnno_L2")+ ggtitle("UMAP Celllabel")
```


## Save the dataset

```{r SaveData}
saveRDS(sce, file =paste0(path,'/output/06_sce_SubClustering.rds'))
```
