---
title: "Subcluster Fibroblast"
subtitle: "06_SC_Fibroblast"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Subcluster Fibroblast

## Setup

### Standard packages

```{r LoadPackages, warning=FALSE}
library(here)
source(here("code", "standard_libraries.R"))
```

### Additional Packages


```{r LoadPackages, warning=FALSE,eval=FALSE}
#Load Packages

suppressPackageStartupMessages({
library(batchelor)
library(bluster)
library(presto)
library(dittoSeq)
library(gridExtra)
})

```

### Set Parameter

```{r Setup}
set.seed(100)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=100)
path <- here::here()
```


### Load Data

```{r LoadDataset}
## RA DATASET
sce <- readRDS(file = paste0(path,'/output/06_sce_SC_Fibroblast.rds'))

```

### Explore Dataset

```{r ShowDim}
#Dimensions of count matrix
dim(sce)
```

https://mperalc.gitlab.io/scRNA-seq_workshop_2021/additional-resources.html## UMAPs coloured by cluster or cell type

### Plot Annotation {.tabset}

#### Plot UMAP of Subclusters 

```{r UMAPCluster}
#cluster
plotReducedDim(sce, paste0("MNN_UMAP_reduced", cell_type),colour_by = "subcluster", text_by="subcluster") +labs(title=paste0("SubCluster ", name),subtitle="Colored by clustering")
```

#### Plot Recalculated UMAP of Subclusters 

```{r UMAPCluster}
#cluster
plotReducedDim(sce, paste0("SC_UMAP_", cell_type),colour_by = "subcluster", text_by="subcluster") +labs(title=paste0("SubCluster ", name),subtitle="Colored by clustering")
```


## DA

### Frequencies of celltypes in clusters per Sample {.tabset}

```{r ClusterInfo}
ClusterInfo <- as.data.frame(colData(sce.sc[,sce.sc$celltype == "Fibroblast"])) %>% 
  group_by(ManualAnno_L2, Sample) %>%
  summarise(cells = n(),.groups = 'drop') 
```

#### Absolute comparison

```{r PlotAbsoluteClustering}

ggplot(data=ClusterInfo, aes(x=Sample,y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="stack") +
  labs(title="Absolute comparison nr Cells in Celltype",subtitle="Per Sample",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())

```

#### Relative comparison

```{r PlotRelativeClustering}

ggplot(data=ClusterInfo,aes(x=Sample, y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="fill") +
  labs(title="Relative comparison nr Cells in celltype",subtitle="Per Sample",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())


```


### Frequencies of celltypes in clusters per Joint Location {.tabset}

```{r ClusterInfo}
ClusterInfo <- as.data.frame(colData(sce.sc[,sce.sc$celltype == "Fibroblast"])) %>% 
  group_by(ManualAnno_L2, Joint.Location) %>%
  summarise(cells = n(),.groups = 'drop') 
```

#### Absolute comparison

```{r PlotAbsoluteClustering}

ggplot(data=ClusterInfo, aes(x=Joint.Location,y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="stack") +
  labs(title="Absolute comparison nr Cells in Celltype",subtitle="Per Joint Location",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())

```

#### Relative comparison

```{r PlotRelativeClustering}

ggplot(data=ClusterInfo,aes(x=Joint.Location, y=cells, fill=ManualAnno_L2)) +
  geom_bar(stat="identity", position="fill") +
  labs(title="Relative comparison nr Cells in celltype",subtitle="Per Joint Location",x="",y="Number of cells")+
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())


```

```{r TestProportionsWilcoxon}

ClusterInfo
wilcox.test()




```




```{r }


Marker <- c("HOXD11","HOXD10","HOXB-AS1","HOXA6","HOXA5","HOXA3","HOXA11-AS","HOXC9","HOXC8","HOXC11","HOXC10","HOXC-AS1","HOXC-AS2","HOXC-AS3","HOXB3","HOXB2")


dittoDotPlot(sce[,sce$celltype == "Fibroblast"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = FALSE )

dittoDotPlot(sce[,sce$celltype == "Fibroblast"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = TRUE )



dittoDotPlot(sce[,sce$celltype == "Edothelial cell"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = FALSE )

dittoDotPlot(sce[,sce$celltype == "Edothelial cell"], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0,max.percent=1,scale = TRUE )



```

## DE

```{r }
-filter genes sc.pp.filter_genes(adata_mono, min_cells=3)
-Function find_de_MAST_RE takes a SingleCellExperiment object as input and runs MAST with RE pipeline. The output of the function is a table (pandas DataFrame in Python) which contains results of the analysis, e.g. log-fold change, p-value and FDR-corrected value for each gene.

find_de_MAST_RE <- function(sce){
    
  # create a MAST object
    print("Dimensions before subsetting:")
    print(dim(sca))
    print("")
    # keep genes that are expressed in more than 10% of all cells
    sca <- sca[freq(sca)>0.1,]
    print("Dimensions after subsetting:")
    print(dim(sca))
    print("")
    # add a column to the data which contains scaled number of genes that are expressed in each cell
    cdr2 <- colSums(assay(sca)>0)
    colData(sca)$ngeneson <- scale(cdr2)
    # store the columns that we are interested in as factors
    label <- factor(colData(sca)$label)
    # set the reference level
    label <- relevel(label,"ctrl")
    colData(sca)$label <- label
    celltype <- factor(colData(sca)$cell_type)
    colData(sca)$celltype <- celltype
    # same for donors (which we need to model random effects)
    replicate <- factor(colData(sca)$replicate)
    colData(sca)$replicate <- replicate
    # create a group per condition-celltype combination
    colData(sca)$group <- paste0(colData(adata_)$label, ".", colData(adata_)$cell_type)
    colData(sca)$group <- factor(colData(sca)$group)
    # define and fit the model
    zlmCond <- zlm(formula = ~ngeneson + group + (1 | replicate), 
                   sca=sca, 
                   method='glmer', 
                   ebayes=F, 
                   strictConvergence=F,
                   fitArgsD=list(nAGQ = 0)) # to speed up calculations
    
    # perform likelihood-ratio test for the condition that we are interested in    
    summaryCond <- summary(zlmCond, doLRT='groupstim.CD14_Monocytes')
    # get the table with log-fold changes and p-values
    summaryDt <- summaryCond$datatable
    result <- merge(summaryDt[contrast=='groupstim.CD14_Monocytes' & component=='H',.(primerid, `Pr(>Chisq)`)], # p-values
                     summaryDt[contrast=='groupstim.CD14_Monocytes' & component=='logFC', .(primerid, coef)],
                     by='primerid') # logFC coefficients
    # MAST uses natural logarithm so we convert the coefficients to log2 base to be comparable to edgeR
    result[,coef:=result[,coef]/log(2)]
    # do multiple testing correction
    result[,FDR:=p.adjust(`Pr(>Chisq)`, 'fdr')]
    result = result[result$FDR<0.01,, drop=F]

    result <- stats::na.omit(as.data.frame(result))
    return(result)
}


-find_de_MAST_RE(adata_mono)
-find de edger()


-plot heatmap
-plot volcano
```







--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------


## Save the dataset

```{r SaveData}
saveRDS(sce, file =paste0(path,'/output/06_sce_SC_Fibroblast.rds'))
```
