---
title: "Subclustering"
subtitle: "07_SubClustering"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Subclustering

## Setup

### Standard packages

```{r LoadPackages, warning=FALSE}
library(here)
source(here("code", "standard_libraries.R"))
```

### Additional Packages

```{r LoadAddPackages, warning=FALSE,eval=FALSE}
#Load Packages

suppressPackageStartupMessages({
library(batchelor)
library(bluster)
library(presto)
library(dittoSeq)
library(gridExtra)
library(intrinsicDimension)
})

```


### Set Parameter

```{r Setup}
set.seed(123)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=123)
path <- here::here()
```


### Load Data

```{r LoadDataset}
## RA DATASET
sce <- readRDS(file = paste0(path,'/output/06_sce_CelltypeAnnotation.rds')) # 06

```


### Explore Dataset

```{r ShowDim}
#Dimensions of count matrix
dim(sce)
```

https://mperalc.gitlab.io/scRNA-seq_workshop_2021/additional-resources.html## UMAPs coloured by cluster or cell type

### Plot Annotation {.tabset}

### Annotation L1

```{r UMAPAnnotationL1}
#cluster
plotReducedDim(sce, "MNN_UMAP_reduced", colour_by="Annotation_L1", text_by="Annotation_L1") +labs(title="UMAP colored by clusters",subtitle = "UMAP of integrated dataset")

```

### Annotation L0

```{r UMAPAnnotationL0}
#cell type
plotReducedDim(sce, "MNN_UMAP_reduced", colour_by="Annotation_L0", text_by="Annotation_L0") +labs(title="UMAP colored by Cell type ",subtitle = "UMAP of integrated dataset")

```

## Subclustering Function

```{r SubclusteringFunction}
# Create Subclustering function

SUBCLUSTER <- function(sce.sc, sc_celltype, clust_param="leiden", k_param=50, merge_order, batchcorr=FALSE){
set.seed(123)

#Subset sce object for ce scelltype
sce.subset <- sce.sc[,sce.sc$celltype == sc_celltype]
    
#Get top HVG for subcluster
dec <- modelGeneVar(sce.subset, block=sce.subset$Sample, density.weights=FALSE,BPPARAM=bpp)
top.hvgs <- getTopHVGs(dec, n=2000)
rowData(sce.subset)$hvg.sc <- rownames(sce.subset) %in% top.hvgs



if(batchcorr==TRUE){
  #Batch correction on subcluster
  sce.Batchelor <- fastMNN(sce.Batchelor, 
                            batch=sce.Batchelor$Sample, 
                            prop.k=0.05,  
                            subset.row = rownames(sce.Batchelor)[rowData(sce.Batchelor)[["hvg.sc"]]], 
                            correct.all=TRUE, merge.order = merge_order, BPPARAM = bpp)
      
  assay(sce.subset, "reconstructed") <- assay(sce.Batchelor, "reconstructed")
  reducedDim(sce.subset, 'MNN_SC') <- reducedDim(sce.Batchelor, 'corrected')
  reducedDim(sce.subset, 'MNN_SC_reduced') <- reducedDim(sce.Batchelor, 'corrected')[,seq_len(15)]
         
  }else{
  #Re run PCA on subset
  sce.subset<-runPCA(sce.subset,exprs_values ="reconstructed", subset_row = rowData(sce.subset)$hvg.sc, name = "MNN_SC")
  reducedDim(sce.subset, 'MNN_SC_reduced') <- reducedDim(sce.subset, 'MNN_SC')[,seq_len(15)]
  
  }

#Get recalculated UMAP on subset
sce.subset <- runUMAP(sce.subset,name = "MNN_UMAP_SC_reduced", dimred = 'MNN_SC_reduced',subset_row=rowData(sce.subset)$hvg.sc)

#Recluster the cells 
set.seed(123)
cluster <- clusterCells(sce.subset, use.dimred = "MNN_SC_reduced", BLUSPARAM = bluster::SNNGraphParam(k= k_param,cluster.fun=clust_param, BPPARAM =bpp))

#Add Subcluster label to sce
colData(sce.subset)$subcluster <- cluster
sce.subset$subcluster <- factor(sce.subset$subcluster)
    
return(sce.subset)
}
```


```{r MergeOrder}
#Set manual merge_order
merge_order <- list(
            list("SynBio_028", "SynBio_049", "SynBio_076", "SynBio_081", "SynBio_087"),#MCP
            list("SynBio_077a","SynBio_093", "SynBio_096","SynBio_098a", "SynBio_127"),#Knee 
            list("SynBio_050", "SynBio_074", "SynBio_077b","SynBio_083","SynBio_084","SynBio_098b"))#Wrist


```


--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

## Subcluster Fibroblast

### Run Subclustering

```{r RunSubclusterFB, warning=FALSE}

#get correct parent label to subcluster
sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.Fibroblast<-SUBCLUSTER(sce, "Fibroblast", clust_param="leiden", k_param=45, merge_order, batchcorr=FALSE)
  
```

```{r dimFB, warning=FALSE}
dim(sce.Fibroblast)
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterFB1C , results='asis'}
plotReducedDim(sce.Fibroblast, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterFB1S , results='asis'}    
plotReducedDim(sce.Fibroblast, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcFB1C , results='asis'}
plotReducedDim(sce.Fibroblast, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcFB1S , results='asis'}
plotReducedDim(sce.Fibroblast, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```

### Plot QC UMAP of Subcluster 

```{r UMAPSubclusterOC1 , results='asis',fig.height=10, fig.width=15}

print(Plot_QC_dimred(sce.Fibroblast,dim="MNN_UMAP_SC_reduced"))
print(Plot_QC_violin(sce.Fibroblast,label="subcluster"))
```

### Remove bad quality subcluster

```{r RmSCFB}
dim(sce.Fibroblast)
sce.Fibroblast <- sce.Fibroblast[,sce.Fibroblast$subcluster != 7]
sce.Fibroblast <- sce.Fibroblast[,sce.Fibroblast$subcluster != 8]
sce.Fibroblast <- sce.Fibroblast[,sce.Fibroblast$subcluster != 9]
dim(sce.Fibroblast)

# Re-Run Subcluster Function
sce.Fibroblast<-SUBCLUSTER(sce.Fibroblast, "Fibroblast", clust_param="leiden", k_param=45, merge_order, batchcorr=FALSE)
```


### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterFB2C , results='asis'}
plotReducedDim(sce.Fibroblast, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterFB2S , results='asis'}    
plotReducedDim(sce.Fibroblast, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcFB2C , results='asis'}
plotReducedDim(sce.Fibroblast, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcFB2S , results='asis'}
plotReducedDim(sce.Fibroblast, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```





```{r , results='asis'}

plotReducedDim(sce.Fibroblast, dimred = "MNN_UMAP_SC_reduced", colour_by = "POSTN" , order_by="POSTN" ,point_alpha=1,point_size=0.5) +scale_colour_gradient(name = "POSTN",low = "lightgrey", high = "red")
plotReducedDim(sce.Fibroblast, dimred = "MNN_UMAP_SC_reduced", colour_by = "CXCL12" , order_by="CXCL12" ,point_alpha=1,point_size=0.5) +scale_colour_gradient(name = "CXCL12",low = "lightgrey", high = "red")
plotReducedDim(sce.Fibroblast, dimred = "MNN_UMAP_SC_reduced", colour_by = "PRG4" , order_by="PRG4" ,point_alpha=1,point_size=0.5) +scale_colour_gradient(name = "PRG4",low = "lightgrey", high = "red")
plotReducedDim(sce.Fibroblast, dimred = "MNN_UMAP_SC_reduced", colour_by = "COMP" , order_by="COMP" ,point_alpha=1,point_size=0.5) +scale_colour_gradient(name = "COMP",low = "lightgrey", high = "red")
plotReducedDim(sce.Fibroblast, dimred = "MNN_UMAP_SC_reduced", colour_by = "MFAP5" , order_by="MFAP5" ,point_alpha=1,point_size=0.5) +scale_colour_gradient(name = "MFAP5",low = "lightgrey", high = "red")
plotReducedDim(sce.Fibroblast, dimred = "MNN_UMAP_SC_reduced", colour_by = "ACTA2" , order_by="ACTA2" ,point_alpha=1,point_size=0.5) +scale_colour_gradient(name = "ACTA2",low = "lightgrey", high = "red")


```


### Subcluster UMAP

```{r FinalUMAPFB}
plotReducedDim(sce.Fibroblast, "MNN_UMAP_SC_reduced",colour_by = "subcluster") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by subcluster")
```

### Save the dataset

```{r SaveDataFB}
saveRDS(sce.Fibroblast, file =paste0(path,'/output/07_sce_SC_Fibroblast.rds'))
rm(sce.Fibroblast)
gc()
```


--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

## Subcluster Macrophage

```{r RunSubclusterMP, warning=FALSE}

sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.Macrophage<-SUBCLUSTER(sce, "Macrophage", clust_param="leiden", k_param=65, merge_order, batchcorr=FALSE)
```


```{r dimMP, warning=FALSE}
dim(sce.Macrophage)
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterMP1C , results='asis'}
plotReducedDim(sce.Macrophage, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterMP1S , results='asis'}    
plotReducedDim(sce.Macrophage, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcMP1C , results='asis'}
plotReducedDim(sce.Macrophage, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcMP1S , results='asis'}
plotReducedDim(sce.Macrophage, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```





```{r UMAPSubclusterOCMP1 , results='asis',fig.height=10, fig.width=15}

Plot_QC_dimred(sce.Macrophage,"MNN_UMAP_SC_reduced")
print(Plot_QC_violin(sce.Macrophage,label="subcluster"))
```

### Remove bad quality subcluster

```{r RmSCMP}
dim(sce.Macrophage)
sce.Macrophage <- sce.Macrophage[,sce.Macrophage$subcluster != 3]
sce.Macrophage <- sce.Macrophage[,sce.Macrophage$subcluster != 4]
dim(sce.Macrophage)

# Re-Run Subcluster Function
sce.Macrophage<-SUBCLUSTER(sce.Macrophage, "Macrophage", clust_param="leiden", k_param=55, merge_order, batchcorr=FALSE)
```


### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterMC2C , results='asis'}
plotReducedDim(sce.Macrophage, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterMC2S , results='asis'}    
plotReducedDim(sce.Macrophage, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcMC2C , results='asis'}
plotReducedDim(sce.Macrophage, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcMC2S , results='asis'}
plotReducedDim(sce.Macrophage, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```


### Subcluster UMAP

```{r FinalUMAPMP}
plotReducedDim(sce.Macrophage, "MNN_UMAP_SC_reduced",colour_by = "subcluster") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by subcluster")
```

### Save the dataset

```{r SaveDataMP}
saveRDS(sce.Macrophage, file =paste0(path,'/output/07_sce_SC_Macrophage.rds'))
rm(sce.Macrophage)
gc()
```

--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

## Subcluster T cell

```{r RunSubclusterTC, warning=FALSE}

sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.Tcell<-SUBCLUSTER(sce, "T cell", clust_param="leiden", k_param=55, merge_order,batchcorr=FALSE) #55
  
```


### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterTC1C , results='asis'}
plotReducedDim(sce.Tcell, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterTC1S , results='asis'}    
plotReducedDim(sce.Tcell, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcTC1C , results='asis'}
plotReducedDim(sce.Tcell, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcTC1S , results='asis'}
plotReducedDim(sce.Tcell, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```


### Plot QC UMAP of Subcluster 

```{r UMAPSubclusterQcTC1 , results='asis',fig.height=10, fig.width=15}

print(Plot_QC_dimred(sce.Tcell,dim="MNN_UMAP_SC_reduced"))
print(Plot_QC_violin(sce.Tcell,label="subcluster"))
```

### Save the dataset

```{r SaveDataTC}
saveRDS(sce.Tcell, file =paste0(path,'/output/07_sce_SC_Tcell.rds'))
rm(sce.Tcell)
gc()
```


--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

## Subcluster Endothelial

```{r RunSubclusterc, warning=FALSE}

sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.Endothelial<-SUBCLUSTER(sce, "Endothelial cell", clust_param="leiden", k_param=50, merge_order, batchcorr=FALSE)
  
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterEC1C , results='asis'}
plotReducedDim(sce.Endothelial, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterEC1S , results='asis'}    
plotReducedDim(sce.Endothelial, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcEC1C , results='asis'}
plotReducedDim(sce.Endothelial, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcEC1S , results='asis'}
plotReducedDim(sce.Endothelial, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```


### Plot QC UMAP of Subcluster

```{r UMAPSubclusterQc1 , results='asis',fig.height=10, fig.width=15}

Plot_QC_dimred(sce.Endothelial,dim="MNN_UMAP_SC_reduced")
Plot_QC_violin(sce.Endothelial,label="subcluster")
```


```{r RmScEc, warning=FALSE}

#remove bad subcluster
dim(sce.Endothelial)
sce.Endothelial <- sce.Endothelial[,sce.Endothelial$subcluster != 3]
dim(sce.Endothelial)

# Run Subcluster Function
sce.Endothelial<-SUBCLUSTER(sce.Endothelial, "Endothelial cell", clust_param="leiden", k_param=39, merge_order, batchcorr=FALSE)
  
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterEC2C , results='asis'}
plotReducedDim(sce.Endothelial, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterEC2S , results='asis'}    
plotReducedDim(sce.Endothelial, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcEC2C , results='asis'}
plotReducedDim(sce.Endothelial, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcEC2S , results='asis'}
plotReducedDim(sce.Endothelial, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```



### Save the dataset

```{r SaveDataEc}
saveRDS(sce.Endothelial, file =paste0(path,'/output/07_sce_SC_Endothelial.rds'))
rm(sce.Endothelial)
gc()
```

--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

## Subcluster SMC

```{r RunSubclusterSMC, warning=FALSE}

sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.SmoothMuscle<-SUBCLUSTER(sce, "Smooth muscle cell", clust_param="leiden", k_param=40, merge_order, batchcorr=FALSE)
  
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterSMC1C , results='asis'}
plotReducedDim(sce.SmoothMuscle, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterSMC1S , results='asis'}    
plotReducedDim(sce.SmoothMuscle, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcSMC1C , results='asis'}
plotReducedDim(sce.SmoothMuscle, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcSMC1S , results='asis'}
plotReducedDim(sce.SmoothMuscle, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```


### Plot QC UMAP of Subcluster 

```{r UMAPSubclusterQcSMC1 , results='asis',fig.height=10, fig.width=15}

print(Plot_QC_dimred(sce.SmoothMuscle,dim="MNN_UMAP_SC_reduced"))
print(Plot_QC_violin(sce.SmoothMuscle,label="subcluster"))

plotReducedDim(sce.SmoothMuscle, dimred = "MNN_UMAP_SC_reduced", colour_by = "ACTA2" , order_by="ACTA2" ,point_alpha=1,point_size=0.5) +scale_colour_gradient(name = "ACTA2",low = "lightgrey", high = "red")

```



```{r RmSCSMC, warning=FALSE}
#remove bad subcluster
dim(sce.SmoothMuscle)
sce.SmoothMuscle <- sce.SmoothMuscle[,sce.SmoothMuscle$subcluster != 2]
dim(sce.SmoothMuscle)

# Run Subcluster Function
#sce.SmoothMuscle<-SUBCLUSTER(sce.SmoothMuscle, "Endothelial cell", clust_param="leiden", k_param=40, merge_order, batchcorr=FALSE)
  
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterSMC2C , results='asis'}
plotReducedDim(sce.SmoothMuscle, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterSMC2S , results='asis'}    
plotReducedDim(sce.SmoothMuscle, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcSMC2C , results='asis'}
plotReducedDim(sce.SmoothMuscle, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcSMC2S , results='asis'}
plotReducedDim(sce.SmoothMuscle, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```

### Plot QC UMAP of Subcluster 

```{r UMAPSubclusterQcSMC2,results='asis',fig.height=10, fig.width=15}

print(Plot_QC_dimred(sce.SmoothMuscle,dim="MNN_UMAP_SC_reduced"))
print(Plot_QC_violin(sce.SmoothMuscle,label="subcluster"))

plotReducedDim(sce.SmoothMuscle, dimred = "MNN_UMAP_SC_reduced", colour_by = "ACTA2" , order_by="ACTA2" ,point_alpha=1,point_size=0.5) +scale_colour_gradient(name = "ACTA2",low = "lightgrey", high = "red")

```




### Save the dataset

```{r SaveDataSMC}
saveRDS(sce.SmoothMuscle, file =paste0(path,'/output/07_sce_SC_SmoothMuscle.rds'))
rm(sce.SmoothMuscle)
gc()
```


--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

## Subcluster Dendrittic Cells

```{r RunSubclusterDC, warning=FALSE}

sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.Dendritic<-SUBCLUSTER(sce, "Dendritic cell", clust_param="leiden", k_param=45, merge_order, batchcorr=FALSE)
  
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterDC1C , results='asis'}
plotReducedDim(sce.Dendritic, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterDC1S , results='asis'}    
plotReducedDim(sce.Dendritic, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcDC1C , results='asis'}
plotReducedDim(sce.Dendritic, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcDC1S , results='asis'}
plotReducedDim(sce.Dendritic, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```


### Plot QC UMAP of Subcluster 

```{r UMAPSubclusterQcDC1 , results='asis',fig.height=10, fig.width=15}

print(Plot_QC_dimred(sce.Dendritic,dim="MNN_UMAP_SC_reduced"))
print(Plot_QC_violin(sce.Dendritic,label="subcluster"))
```





### Save the dataset

```{r SaveDataDC}
saveRDS(sce.Dendritic, file =paste0(path,'/output/07_sce_SC_Dendritic.rds'))
rm(sce.Dendritic)
gc()
```

--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

## Subcluster Mast Cells

```{r RunSubclusterMC, warning=FALSE}

sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.Mast<-SUBCLUSTER(sce, "Mast cell", clust_param="leiden", k_param=20, merge_order, batchcorr=FALSE)
  
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterMC1C , results='asis'}
plotReducedDim(sce.Mast, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterMC1S , results='asis'}    
plotReducedDim(sce.Mast, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcMC1C , results='asis'}
plotReducedDim(sce.Mast, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcMC1S , results='asis'}
plotReducedDim(sce.Mast, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```

### Plot QC UMAP of Subclusters 

```{r UMAPSubclusterQcMC1 , results='asis',fig.height=10, fig.width=15}

print(Plot_QC_dimred(sce.Mast,dim="MNN_UMAP_SC_reduced"))
print(Plot_QC_violin(sce.Mast,label="subcluster"))

plotReducedDim(sce.Mast, dimred = "MNN_UMAP_SC_reduced", colour_by = "TPSAB1" , order_by="TPSAB1" ,point_alpha=1,point_size=0.5) +scale_colour_gradient(name = "TPSAB1",low = "lightgrey", high = "red")
```



```{r RmSCMC, warning=FALSE}
#remove bad subcluster
dim(sce.Mast)
sce.Mast <- sce.Mast[,sce.Mast$subcluster == 2]
dim(sce.Mast)
```


### Save the dataset

```{r SaveDataMC}
saveRDS(sce.Mast, file =paste0(path,'/output/07_sce_SC_Mast.rds'))
rm(sce.Mast)
gc()
```


--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------

## Subcluster B cell

```{r RunSubclusterBC, warning=FALSE}

sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.Bcell<-SUBCLUSTER(sce, "B cell", clust_param="leiden", k_param=50, merge_order, batchcorr=FALSE)
  
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterBC1C , results='asis'}
plotReducedDim(sce.Bcell, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterBC1S , results='asis'}    
plotReducedDim(sce.Bcell, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcBC1C , results='asis'}
plotReducedDim(sce.Bcell, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcBC1S , results='asis'}
plotReducedDim(sce.Bcell, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```



### Plot QC UMAP of Subclusters 

```{r UMAPSubclusterQcBC1 , results='asis',fig.height=10, fig.width=15}

print(Plot_QC_dimred(sce.Bcell,dim="MNN_UMAP_SC_reduced"))
print(Plot_QC_violin(sce.Bcell,label="subcluster"))

```

### Save the dataset

```{r SaveDataC}
saveRDS(sce.Bcell, file =paste0(path,'/output/07_sce_SC_Bcell.rds'))
rm(sce.Bcell)
gc()
```


--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------


## Subcluster Plasma cell

```{r RunSubclusterPC, warning=FALSE}

sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.Plasma<-SUBCLUSTER(sce, "Plasma", clust_param="leiden", k_param=50, merge_order, batchcorr=FALSE)
  
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterPC1C , results='asis'}
plotReducedDim(sce.Plasma, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterPC1S , results='asis'}    
plotReducedDim(sce.Plasma, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcPC1C , results='asis'}
plotReducedDim(sce.Plasma, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcPC1S , results='asis'}
plotReducedDim(sce.Plasma, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```


### Plot QC UMAP of Subclusters 

```{r UMAPSubclusterQcPC1 , results='asis',fig.height=10, fig.width=15}

print(Plot_QC_dimred(sce.Plasma,dim="MNN_UMAP_SC_reduced"))
print(Plot_QC_violin(sce.Plasma,label="subcluster"))

```


### Save the dataset

```{r SaveDataPc}
saveRDS(sce.Plasma, file =paste0(path,'/output/07_sce_SC_Plasma.rds'))
rm(sce.Plasma)
gc()
```



--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------



## Subcluster Neutrophil

```{r RunSubclusterNP, warning=FALSE}

sce$celltype <- colData(sce)$Annotation_L1

# Run Subcluster Function
sce.Neutrophil<-SUBCLUSTER(sce, "Neutrophil", clust_param="leiden", k_param=50, merge_order, batchcorr=FALSE)
  
```

### Plot UMAP of Subclusters  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterNP1C , results='asis'}
plotReducedDim(sce.Neutrophil, "MNN_UMAP_reduced",colour_by = "subcluster", text_by="subcluster") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterNP1S , results='asis'}    
plotReducedDim(sce.Neutrophil, "MNN_UMAP_reduced",colour_by = "Sample") +
          labs(title=paste0("UMAP of Subcluster"),subtitle="Colored by Sample")
```


### Plot Recalculated UMAP of Subcluster  {.tabset}

#### Colored by Clustering
```{r UMAPSubclusterRecalcNP1C , results='asis'}
plotReducedDim(sce.Neutrophil, "MNN_UMAP_SC_reduced",colour_by = "subcluster",text_by="subcluster") +labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by clustering")
```

#### Colored by Sample
```{r UMAPSubclusterRecalcNP1S , results='asis'}
plotReducedDim(sce.Neutrophil, "MNN_UMAP_SC_reduced",colour_by = "Sample") +
          labs(title=paste0("Recalculated UMAP for Subcluster"),subtitle="Colored by Sample")

```

### Plot QC UMAP of Subclusters 

```{r UMAPSubclusterQcNP1 , results='asis',fig.height=10, fig.width=15}

print(Plot_QC_dimred(sce.Neutrophil,dim="MNN_UMAP_SC_reduced"))
print(Plot_QC_violin(sce.Neutrophil,label="subcluster"))

```

### Save the dataset

```{r SaveDataNP}
saveRDS(sce.Neutrophil, file =paste0(path,'/output/07_sce_SC_Neutrophil.rds'))
rm(sce.Neutrophil)
gc()
```





