---
title: "Clustering"
subtitle: "10_Plotting"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Plotting

## Setup

### Standard packages

```{r LoadPackages, warning=FALSE}
library(here)
source(here("code", "standard_libraries.R"))
```

### Additional Packages

```{r LoadAddPackages, warning=FALSE}
#Load Packages

suppressPackageStartupMessages({
library(gridExtra)
library(dittoSeq)
library(ggpubr)
})

```


### Set Parameter

```{r Setup}
set.seed(123)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=123)
path <- here::here()
```


-------------------------------------------------------------------------------------------------------------------------------------------------------------
```{r}
# RA DATASET
sce <- readRDS(file = paste0(path,'/output/03_sce_Integration_Batchelor.rds'))#03_sce_Integration_Batchelor.rds
#sce <- readRDS(file = paste0(path,'/output/03_sce_Integration_Harmony.rds'))#03_sce_Integration_Batchelor.rds

#Change to genesymbol whenever possible
rowData(sce)$UniqID <-rownames(sce)
rownames(sce) <- uniquifyFeatureNames(rowData(sce)$ENSEMBL, rowData(sce)$Symbol)
```

```{r}
#Heatmap of Marker Gened used for Annotation
#Endo "PECAM1", "CDH5"
#Myeloid "CD14","FCGR3A"
#Bcell "CD19","MS4A1","CD79A"
#Plasma "SDC1","MZB1"
#Tcell "CD3E", "CD3G"
#Mast "KIT","IL1RL1","MS4A2"
#Fibroblast "PDGFRB","DCN","COL1A1"
#Neutrophil "CSF3R","S100A8","TREM1"
#pDC
#SMC
markr<-c("CDH5","CD14","CD19","SDC1","CD3E","KIT","COL1A1","CSF3R","ACTA2","CLEC4C","UPK3A","CLEC9A","CCL22","PROX1","LYVE1")

```


```{r}
# RA DATASET
for(i in markr){
  print(plotReducedDim(sce, dimred="MNN_UMAP_reduced", colour_by = i, order_by=i,point_alpha=1,point_size=0.5 )+scale_colour_gradient(name = i,low = "lightgrey", high = "red"))#harmony_reduced_UMAP
  }


```

-------------------------------------------------------------------------------------------------------------------------------------------------------------

## UMAP

### Load Data

```{r}
# RA DATASET
sce <- readRDS(file = paste0(path,'/output/05_sce_CelltypeAnnotation.rds'))

```

```{r SampleMetricsTranspose}
  umap_dt = data.table(
    cell_id   = rownames(umap_x),
    idv_umap1 = umap_x[, 1],
    idv_umap2 = umap_x[, 2]
    )

nice_cols   = c(
    "#DC050C", "#FB8072", "#1965B0", "#7BAFDE", "#882E72", "#B17BA6",
    "#FF7F00", "#FDB462", "#E7298A", "#E78AC3", "#33A02C", "#B2DF8A",
    "#55A1B1", "#8DD3C7", "#A6761D", "#E6AB02", "#7570B3", "#BEAED4",
    "#666666", "#999999", "#aa8282", "#d4b7b7", "#8600bf", "#ba5ce3",
    "#808000", "#aeae5c", "#1e90ff", "#00bfff", "#56ff0d", "#ffff00"
    )

sce$MNN_UMAP_reduced[,1]
sce$MNN_UMAP_reduced[,2]

plotReducedDim(sce, dimred="MNN_UMAP_reduced", colour_by="label",order_by="label",point_alpha=1,point_size=1)
 
ggplot(as.data.frame(colData(sce))) + 
    aes(x = reducedDim(sce,"MNN_UMAP_reduced")[,1], y = reducedDim(sce,"MNN_UMAP_reduced")[,2], colour = label) + 
    geom_point( size=1.5 ) + 
    scale_colour_manual( values=nice_cols[seq.int(length(unique(sce$label)))] ) + theme( axis.text=element_blank(), legend.position='right' ) +
    labs( x='UMAP1', y='UMAP2' )+ theme_classic()



```

-------------------------------------------------------------------------------------------------------------------------------------------------------------

## HOX genes plots

### Load Data

```{r}
# RA DATASET
sce <- readRDS(file = paste0(path,'/output/05_sce_CelltypeAnnotation.rds'))
Marker <- c("HOXD11","HOXD13","HOXD10","HOXB-AS1","HOXA6","HOXA5","HOXA3","HOXA11-AS","HOXC9","HOXC8","HOXC11","HOXC10","HOXC-AS1","HOXC-AS2","HOXC-AS3","HOXB3","HOXB2")

Marker <- c("HOXD13","HOXD11","HOXD10","HOXD-AS2","HOXD9","HOXD8","HOXD3","HOXD4","HOXD1","HOXA1","HOXA2","HOXA3","HOXA-AS2","HOXA4","HOXA-AS3","HOXA5","HOXA6","HOXA7","HOXA9","HOXA10-AS","HOXA10","HOXA11","HOXA11-AS","HOXA13","HOXC13","HOXC11","HOXC-AS3" ,"HOXC10","HOXC6","HOXC-AS2","HOXC9","HOXC-AS1","HOXC8","HOXC4","HOXC5","HOXB2","HOXB-AS1","HOXB3","HOXB-AS3","HOXB4","HOXB5","HOXB6","HOXB7","HOXB8","HOXB9","HOXB13")

```

### Dotplot

```{r }

Dotplot_facet <- function(sce, Marker, label, condition) {
  p <- list() 
  for (c_type in label){

    p[[c_type]]<-dittoDotPlot(sce[,sce$celltype == c_type], vars = Marker , group.by = "Joint.Location",min.color="lightgrey", max.color="blue",min.percent=0.1,max.percent=1,scale = FALSE,max=2 )+ theme(axis.title.x=element_blank())+
      coord_flip()+ ggtitle(c_type)+ theme(plot.title = element_text(hjust = 0.5))
}
  return(p) 
}

```


```{r , fig.width=15}

p<- Dotplot_facet(sce, Marker, unique(sce$label), condition="Joint.Location")

grid.arrange(p$`Edothelial cell` + theme(legend.position="none"), 
             p$Fibroblast + theme(legend.position="none",axis.line.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank()), 
             p$Myeloid + theme(legend.position="none",axis.line.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank()), 
             p$Plasma + theme(legend.position="none",axis.line.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank()), 
             p$`T cell` + theme(legend.position="none",axis.line.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank()), 
             p$`B cell` + theme(legend.position="none",axis.line.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank()), 
             p$Neutrophil + theme(axis.line.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank()), 
             ncol = 7, widths = c(1.5/8, 1/8, 1/8, 1/8, 1/8, 1/8, 1.5/8))


#p[[1]] <- p[[1]]+ theme(legend.position="none")
#p[seq(2,length(p)-1)]<-lapply(p[seq(2,length(p)-1)], function(x) x<- x + theme(legend.position="none",axis.line.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank()))
#p[[length(p)]] <-p[[length(p)]]+ theme(axis.line.y=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank())

#do.call("grid.arrange", c(p, ncol=length(p), widths = c(1.5/8, 1/8, 1/8, 1/8, 1/8, 1/8, 1.5/8)))


#add lines between plots
x = c(1.5/8, 1.5/8, 2.5/8, 2.5/8, 3.5/8, 3.5/8, 4.5/8, 4.5/8, 5.5/8, 5.5/8, 6.5/8, 6.5/8)
y = rep(c(0.1, 0.92),times=6)
id = c(1,1,2,2,3,3,4,4,5,5,6,6)
grid::grid.polygon(x,y,id)
```

### Histogramm

```{r }

Histplot_facet <- function(sce, Marker, condition) {
  p <- list() 
  for (m in Marker){
    p[[m]] <- dittoPlot(sce, m, group.by = condition, split.by = 'label' ,split.ncol = 7,max=2)+
      theme(axis.title.x=element_blank(),plot.title=element_blank(),legend.position="none",strip.background = element_blank(),strip.text.x = element_blank(), axis.text.x=element_blank())
      
}
  return(p) 
}


```

```{r }

p<- Histplot_facet(sce, Marker, condition="Joint.Location")


do.call("grid.arrange", c(p, nrow=length(p)))
#ggarrange(plotlist=p,
          
```

-------------------------------------------------------------------------------------------------------------------------------------------------------------

## Sample distances

A useful first step in an RNA-seq analysis is often to assess overall similarity between samples: Which samples are similar to each other, which are different? Does this fit to the expectation from the experiment’s design? We use the R function dist to calculate the Euclidean distance between samples.

```{r SampleMetricsTranspose}
library("pheatmap")
library("RColorBrewer")
library("PoiClaClu")

plot_dist_heatmap <- function(sce, sampleDists) {
  
  sampleDistMatrix <- as.matrix(sampleDists)
  rownames(sampleDistMatrix) <- unique(paste( sce$Joint.Location, sce$Sample, sep = ":" ))
  colnames(sampleDistMatrix) <- NULL
  colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
  dist_heatmap<-pheatmap(sampleDistMatrix,
           clustering_distance_rows = sampleDists,
           clustering_distance_cols = sampleDists,
           col = colors)

  return(dist_heatmap) 
}

```



```{r VariableMetrics}

sce.subset <- sce[rowData(sce)$is_hvg== "TRUE",sce$celltype == "Fibroblast"]
sce.subset.aggr <- aggregateAcrossCells(sce.subset,  ids = sce.subset$Sample, statistics = "mean",use.assay.type = "logcounts")

sampleDists <- dist(t(assay(sce.subset.aggr)))
print(plot_dist_heatmap(sce.subset,sampleDists))

poisd <- PoissonDistance(t(assay(sce.subset.aggr)))
print(plot_dist_heatmap(sce.subset,poisd$dd))


```
