---
title: "Subcluster Annotation"
subtitle: "10_CellChat"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Cell-Cell Communication

## Setup

### Load packages

```{r LoadPackages, warning=FALSE}
#Standard Packages
library(here)
source(here("code", "standard_libraries.R"))

#Additional Packages
suppressPackageStartupMessages({
library(CellChat)
library(patchwork)
})

```

### Set Parameter

```{r Setup}
set.seed(100)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=100)
path <- here::here()
```



### Load Data

```{r LoadDataset}
## RA DATASET
sce <- readRDS(file = paste0(path,'/output/07_sce_SC_Merge.rds'))

```

### Explore Dataset

```{r ShowDim}
#Dimensions of count matrix
dim(sce)
```


### Plot Annotation {.tabset}

#### Annotation L2

```{r UMAPCluster}
#cluster
plotReducedDim(sce, "MNN_merge_reduced_UMAP", colour_by="Annotation_L2") +labs(title="UMAP colored by clusters",subtitle = "UMAP of integrated dataset")

```

#### Annotation L1

```{r UMAPCluster}
#cluster
plotReducedDim(sce, "MNN_merge_reduced_UMAP", colour_by="Annotation_L2", text_by="Annotation_L2") +labs(title="UMAP colored by clusters",subtitle = "UMAP of integrated dataset")

```

#### Annotation L0

```{r UMAPAnnotation}
#cell type
plotReducedDim(sce, "MNN_merge_reduced_UMAP", colour_by="Annotation_L0", text_by="Annotation_L0") +labs(title="UMAP colored by Cell type ",subtitle = "UMAP of integrated dataset")

```


## Setup

```{r}

#Create Annotation
sce$CellChatAnno <- sce$Annotation_L2
unique(sce$CellChatAnno)

sce$CellChatAnno[sce$Annotation_L2 %in% c("CytotoxCD4 TC","Memory TC","Naive TC","CytotoxCD8 TC","NKTcell")] <- "T cell"
sce$CellChatAnno[sce$Annotation_L2 %in% c("TREM2+ MP","FOLR2+ MP")] <- "Tissue-resident Macrophage"
sce$CellChatAnno[sce$Annotation_L2 %in% c("S100A12+ MP","CD48+ MP")] <- "Macrophage"
sce$CellChatAnno[sce$Annotation_L2 %in% c("cDC","pDC")] <- "Dendritic cell"
sce$CellChatAnno[sce$Annotation_L2 %in% c("Arteriolar EC","Venular EC","Capillary EC")] <- "Endothelial"

#Create a CellChat object
cellChat.MCP <- createCellChat(object = sce[,sce$Joint.Location == "MCP"], group.by = "CellChatAnno")
cellChat.Knee <- createCellChat(object = sce[,sce$Joint.Location == "Knee"], group.by = "CellChatAnno")
cellChat.Wrist <- createCellChat(object = sce[,sce$Joint.Location == "Wrist"], group.by = "CellChatAnno")

#unique(cellchat@meta$CellChatAnno)
```

## Set the ligand-receptor interaction database

```{r}
CellChatDB <- CellChatDB.human 
showDatabaseCategory(CellChatDB)

# use all CellChatDB for cell-cell communication analysis
#CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling", key = "annotation") # use Secreted Signaling
CellChatDB.use <- subsetDB(CellChatDB) # use all Signaling


# use all CellChatDB except for "Non-protein Signaling" for cell-cell communication analysis
# CellChatDB.use <- subsetDB(CellChatDB)

# set the used database in the object
cellChat.MCP@DB <- CellChatDB.use
cellChat.Knee@DB <- CellChatDB.use
cellChat.Wrist@DB <- CellChatDB.use
```

## Preprocessing expression data for cell-cell communication analysis

```{r}
# subset the expression data of signaling genes for saving computation cost
cellChat.MCP <- subsetData(cellChat.MCP) # This step is necessary even if using the whole database
cellChat.Knee <- subsetData(cellChat.Knee)
cellChat.Wrist <- subsetData(cellChat.Wrist)

future::plan("multisession", workers = 4) # do parallel

cellChat.MCP <- identifyOverExpressedGenes(cellChat.MCP)
cellChat.Knee <- identifyOverExpressedGenes(cellChat.Knee)
cellChat.Wrist <- identifyOverExpressedGenes(cellChat.Wrist)

cellChat.MCP <- identifyOverExpressedInteractions(cellChat.MCP)
cellChat.Knee <- identifyOverExpressedInteractions(cellChat.Knee)
cellChat.Wrist <- identifyOverExpressedInteractions(cellChat.Wrist)

#cellChat.MCP <- projectData(cellChat.MCP, PPI.human)
#cellChat.Knee <- projectData(cellChat.Knee, PPI.human)
#cellChat.Wrist <- projectData(cellChat.Wrist, PPI.human)
```


## Inference of cell-cell communication network

```{r}
cellChat.MCP <- computeCommunProb(cellChat.MCP, type = "triMean")
cellChat.Knee <- computeCommunProb(cellChat.Knee, type = "triMean")
cellChat.Wrist <- computeCommunProb(cellChat.Wrist, type = "triMean")
```

```{r}
cellChat.MCP <- filterCommunication(cellChat.MCP, min.cells = 10)
cellChat.Knee <- filterCommunication(cellChat.Knee, min.cells = 10)
cellChat.Wrist <- filterCommunication(cellChat.Wrist, min.cells = 10)
```

## Infer the cell-cell communication at a signaling pathway level

```{r}
cellChat.MCP <- computeCommunProbPathway(cellChat.MCP)
cellChat.Knee <- computeCommunProbPathway(cellChat.Knee)
cellChat.Wrist <- computeCommunProbPathway(cellChat.Wrist)

cellChat.MCP <- aggregateNet(cellChat.MCP)
cellChat.Knee <- aggregateNet(cellChat.Knee)
cellChat.Wrist <- aggregateNet(cellChat.Wrist)

cellChat.MCP<-netAnalysis_computeCentrality(cellChat.MCP)
cellChat.Knee<-netAnalysis_computeCentrality(cellChat.Knee)
cellChat.Wrist<-netAnalysis_computeCentrality(cellChat.Wrist)
```

## Save the datasets

```{r SaveData}
saveRDS(cellChat.MCP, file =paste0(path,'/output/10_sce_CellChat_MCP.rds'))
saveRDS(cellChat.Knee, file =paste0(path,'/output/10_sce_CellChat_Knee.rds'))
saveRDS(cellChat.Wrist, file =paste0(path,'/output/10_sce_CellChat_Wrist.rds'))
```



## Load CellChat object of each dataset and merge them together

### Load CellChat
```{r}
cellChat.MCP <- readRDS(file =paste0(path,'/output/10_sce_CellChat_MCP.rds'))
cellChat.Knee <- readRDS(file =paste0(path,'/output/10_sce_CellChat_Knee.rds'))
cellChat.Wrist <- readRDS(file =paste0(path,'/output/10_sce_CellChat_Wrist.rds'))
```


### Merge CellChat objects

```{r}
cellChat.MCP<-updateCellChat(cellChat.MCP)
cellChat.Knee<-updateCellChat(cellChat.Knee)
cellChat.Wrist<-updateCellChat(cellChat.Wrist)

object.list <- list(MCP = cellChat.MCP, Knee = cellChat.Knee, Wrist = cellChat.Wrist)#list(MCP = cellChat.MCP, Knee = cellChat.Knee, Wrist = cellChat.Wrist)
cellchat <- mergeCellChat(object.list, add.names = names(object.list),cell.prefix = TRUE)
cellchat
```


## Comparison Analysis

### Compare the total number of interactions and interaction strength
```{r}
gg1 <- compareInteractions(cellchat, show.legend = F, group = c("MCP","Knee","Wrist"),color.use =c("#FF5053","#4284FF","#00B600"), size.text = 15)+scale_y_continuous(expand = c(0, 0), limits = c(0, 15900))
gg2 <- compareInteractions(cellchat, show.legend = F, group = c("MCP","Knee","Wrist"),color.use =c("#FF5053","#4284FF","#00B600"), measure = "weight", size.text = 15)+scale_y_continuous(expand = c(0, 0), limits = c(0, 340))
gg1 + gg2
```


```{r}
# Extract differential interactions
interaction_diff <- cellchat@net$compare$result
head(interaction_diff)
```


### Differential number of interactions or interaction strength among coarse cell types
```{r}
#levels(cellchat@idents)
#group.cellType <- c(rep("FIB", 4), rep("DC", 4), rep("TC", 4))
#group.cellType <- factor(group.cellType, levels = c("FIB", "DC", "TC"))
#object.list <- lapply(object.list, function(x) {mergeInteractions(x, group.cellType)})
#cellchat <- mergeCellChat(object.list, add.names = names(object.list))



#par(mfrow = c(1,2), xpd=TRUE)
#for (i in 1:length(object.list)) {
  #netVisual_circle(object.list[[i]]@net$count.merged, weight.scale = T, label.edge= T, edge.weight.max = weight.max[3], edge.width.max = 12, title.name = paste0("Number of interactions - ", names(object.list)[i]))
#}
```




### Compare the number of interactions and interaction strength among different cell populations
```{r}
#sources.use =c("CHI3L2+ FIB","CXCL12+ FIB","MFAP5+ FIB","POSTN+ FIB","PRG4+ FIB")
#targets.use = c("FOLR2+ MP","TREM2+ MP")

#KneeVsMCP
gg1 <- netVisual_heatmap(cellchat,comparison = c("Knee","MCP"),title.name = "Knee Vs MCP")
gg2 <- netVisual_heatmap(cellchat,comparison = c("Knee","MCP"), measure = "weight",title.name = "Knee Vs MCP")
gg1 + gg2

#KneeVsWrist
gg1 <- netVisual_heatmap(cellchat,comparison = c("Knee","Wrist"),title.name = "Knee Vs Wrist")
gg2 <- netVisual_heatmap(cellchat,comparison = c("Knee","Wrist"), measure = "weight",title.name = "Knee Vs Wrist")
gg1 + gg2

#WRISTVsMCP
gg1 <- netVisual_heatmap(cellchat,comparison = c("MCP","Wrist"),title.name = "MCP Vs Wrist")
gg2 <- netVisual_heatmap(cellchat, comparison = c("MCP","Wrist"), measure = "weight",title.name = "MCP Vs Wrist")
gg1 + gg2
```


### Identify cell populations with significant changes in sending or receiving signals

```{r}

num.link <- sapply(object.list, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg <- list()
for (i in 1:length(object.list)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(object.list[[i]], title = names(object.list)[i], weight.MinMax = weight.MinMax)
}
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
patchwork::wrap_plots(plots = gg)
```

```{r}
#POSTN+ FIB
gg1 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 1), idents.use = "PRG4+ FIB", signaling.exclude = c("MIF"))
gg2 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 3), idents.use = "PRG4+ FIB", signaling.exclude = c("MIF"))
gg3 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(1, 3), idents.use = "PRG4+ FIB", signaling.exclude = c("MIF"))


patchwork::wrap_plots(plots = list(gg1,gg2,gg3))


#Tissue-resident Macrophage 
gg1 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 1), idents.use = "Tissue-resident Macrophage", signaling.exclude = c("MIF"))
gg2 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 3), idents.use = "Tissue-resident Macrophage", signaling.exclude = c("MIF"))
gg3 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(1, 3), idents.use = "Tissue-resident Macrophage", signaling.exclude = c("MIF"))


patchwork::wrap_plots(plots = list(gg1,gg2,gg3))


#CHI3L2+ FIB
gg1 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 1), idents.use = "CHI3L2+ FIB", signaling.exclude = c("MIF"))
gg2 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 3), idents.use = "CHI3L2+ FIB", signaling.exclude = c("MIF"))
gg3 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(1, 3), idents.use = "CHI3L2+ FIB", signaling.exclude = c("MIF"))


#patchwork::wrap_plots(plots = list(gg1,gg2,gg3))


#CHI3L2+ FIB
gg1 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 1), idents.use = "Neutrophil", signaling.exclude = c("MIF"))
gg2 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 3), idents.use = "Neutrophil", signaling.exclude = c("MIF"))
gg3 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(1, 3), idents.use = "Neutrophil", signaling.exclude = c("MIF"))


patchwork::wrap_plots(plots = list(gg1,gg2,gg3))

```

### Identify cell populations with significant changes in sending or receiving signals

```{r}
pathways.show <- c("SEMA3") 
par(mfrow = c(1,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
  ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))
```


### Identify cell populations with significant changes in sending or receiving signals

```{r}
gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE,comparison = c(1, 2))
gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE,comparison = c(1, 2))

gg1 + gg2


gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE,comparison = c(1, 3))
gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE,comparison = c(1, 3))

gg1 + gg2


gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE,comparison = c(3, 2))
gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE,comparison = c(3, 2))

gg1 + gg2

rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE,comparison = c(1, 2, 3))

rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial"), targets.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial"), stacked = F, do.stat = TRUE,comparison = c(1, 2, 3))





#######
rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial"), targets.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial"), stacked = F, do.stat = TRUE,comparison = c(1, 2, 3),signaling=c("SEMA3","GAP","ncWNT","WNT","MMP","PTN","CXCL","FGF","EGF","GAP","ANGPTL","SELE","NOTCH","VEGF","TGFb","IL16","ADGRL","PDGF","MIF","IL6","VEGEF","ICAM"))


#Check sig

targets = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial")

gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = targets, targets.use = targets, stacked = T, do.stat = TRUE,comparison = c(1, 2))
gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = targets, targets.use = targets, stacked = F, do.stat = TRUE,comparison = c(1, 2))

gg1 + gg2


gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = targets, targets.use = targets, stacked = T, do.stat = TRUE,comparison = c(1, 3))
gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = targets, targets.use = targets, stacked = F, do.stat = TRUE,comparison = c(1, 3))

gg1 + gg2


gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = targets, targets.use = targets, stacked = T, do.stat = TRUE,comparison = c(3, 2))
gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = targets, targets.use = targets, stacked = F, do.stat = TRUE,comparison = c(3, 2))

gg1 + gg2



#----IL 6


pathways.show <- c("IL6") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))


netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("Endothelial"), targets.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), angle.x = 45,remove.isolate = FALSE,signaling=c("IL6"))


#----WNT


pathways.show <- c("WNT") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))


netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("Tissue-resident Macrophage"), targets.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), angle.x = 45,remove.isolate = FALSE,signaling=c("WNT"))


#----VEGF


pathways.show <- c("VEGF") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))



netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Tissue-resident Macrophage"), targets.use = c("Endothelial"), angle.x = 45,remove.isolate = FALSE,signaling=c("VEGF"))

#####


#----MMP


pathways.show <- c("MMP") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i])
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))



netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Endothelial"), targets.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Endothelial"), angle.x = 45,remove.isolate = FALSE,signaling=c("VEGF"))


#----FGF
pathways.show <- c("FGF") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))


netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), targets.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Endothelial"), angle.x = 45,remove.isolate = FALSE,signaling=c("FGF"))




#----FGF
pathways.show <- c("TGFb") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))


netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), targets.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Endothelial"), angle.x = 45,remove.isolate = FALSE,signaling=c("FGF"))


#---PDGF

netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("POSTN+ FIB","PRG4+ FIB","CHI3L2+ FIB"), targets.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), angle.x = 45,remove.isolate = FALSE,signaling=c("PDGF"))







netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("Tissue-resident Macrophage","PRG4+ FIB","CHI3L2+ FIB"), targets.use = c("Tissue-resident Macrophage","PRG4+ FIB","CHI3L2+ FIB"), angle.x = 45,remove.isolate = FALSE,signaling=c("ANGPTL"))



```


```{r}
###PLOTS THIS IS THE ONEEEE



gg1 <- netVisual_heatmap(cellchat,comparison = c("Knee","MCP"),title.name = "Knee Vs MCP",cluster.rows = TRUE,cluster.cols = TRUE)

gg2 <- netVisual_heatmap(cellchat,comparison = c("Knee","Wrist"),title.name = "Knee Vs Wrist",cluster.rows = TRUE,cluster.cols = TRUE)

gg3 <- netVisual_heatmap(cellchat,comparison = c("MCP","Wrist"),title.name = "MCP Vs Wrist",cluster.rows = TRUE,cluster.cols = TRUE)#,color.use = rep("grey",15)

gg1 + gg2 + gg3



###CHECK SIG VALUE!!

a<-rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE,comparison = c(1, 3),return.data = TRUE)
a$signaling.contribution%>% filter(name == "COLLAGEN")
------





##ALL

rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial","Macrophage"), targets.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial","Macrophage"), stacked = F, do.stat = TRUE,comparison = c(1, 2, 3),signaling=c("SEMA3","GAP","ncWNT","WNT","MMP","PTN","CXCL","FGF","EGF","GAP","ANGPTL","SELE","NOTCH","VEGF","TGFb","ADGRL","PDGF","MIF","IL6","VEGEF"))



###IL6

pathways.show <- c("IL6") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))


netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("Endothelial"), targets.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), angle.x = 45,remove.isolate = FALSE,signaling=c("IL6"))

plotGeneExpression(cellchat, signaling = "IL6", split.by = "datasets", colors.ggplot = T, type = "violin")



#Circle plots
pathways.show <- c("IL6") 
weight.max <- getMaxWeight(object.list, slot.name = c("netP"), attribute = pathways.show) # control the edge weights across different datasets
par(mfrow = c(1,1), xpd=TRUE)
for (i in 1:length(object.list)) {
    netVisual_aggregate(object.list[[i]], signaling = pathways.show, top = 0.1,layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object.list)[i]))
}




##Neutrophil Target
#signaling=c("PECAM1","NOTCH","SAA","SEMA7","SELE","EPHB","COLLAGEN","IL1","ANGPTL","ADGRE","GALECTIN","PLAU","IGF","SPP1","ANNEXIN","CXCL","CTSG","SELPLG","COMPLEENT","ICAM","THY1")
signaling=c("COLLAGEN","ADGRE","ANNEXIN")
rankNet(cellchat, mode = "comparison", measure = "weight", targets.use = c("Neutrophil"), sources.use = NULL,stacked = F, do.stat = TRUE,comparison = c(1, 2, 3),signaling=signaling)


netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("Tissue-resident Macrophage","Endothelial","PRG4+ FIB","CHI3L2+ FIB","CXCL12+ FIB","Macrophage","MFAP5+ FIB","POSTN+ FIB"), targets.use = c("Neutrophil"), angle.x = 45,remove.isolate = FALSE,signaling=c("ADGRE"),dot.size.max=4)






###SPATIAL CORR
rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage","PRG4+ FIB","CHI3L2+ FIB"), targets.use = c("Tissue-resident Macrophage","PRG4+ FIB","CHI3L2+ FIB"), stacked = F, do.stat = TRUE,comparison = c(1, 2, 3))


##MIF
netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("Tissue-resident Macrophage","PRG4+ FIB","CHI3L2+ FIB"), targets.use = c("Tissue-resident Macrophage","PRG4+ FIB","CHI3L2+ FIB"), angle.x = 45,remove.isolate = FALSE,signaling=c("MIF"),dot.size.max  =4)


##ANGPTL

rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage","PRG4+ FIB","CHI3L2+ FIB"), targets.use = c("Tissue-resident Macrophage","PRG4+ FIB","CHI3L2+ FIB"), stacked = F, do.stat = TRUE,comparison = c(1, 2, 3),signaling=c("ANGPTL"))

netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("PRG4+ FIB","CHI3L2+ FIB"), targets.use = c("Tissue-resident Macrophage","PRG4+ FIB","CHI3L2+ FIB"), angle.x = 45,remove.isolate = FALSE,signaling=c("ANGPTL"),dot.size.min =4)




###WNT

pathways.show <- c("WNT") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))


netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("Tissue-resident Macrophage"), targets.use = c("PRG4+ FIB","CHI3L2+ FIB","POSTN+ FIB","CXCL12+ FIB","MFAP5+ FIB"), angle.x = 40,remove.isolate = FALSE,signaling=c("WNT"),dot.size.min =4)




plotGeneExpression(cellchat, signaling = "IL6", split.by = "datasets", colors.ggplot = T, type = "violin")



#Circle plots
pathways.show <- c("WNT") 
weight.max <- getMaxWeight(object.list, slot.name = c("netP"), attribute = pathways.show) # control the edge weights across different datasets
par(mfrow = c(1,1), xpd=TRUE)
for (i in 1:length(object.list)) {
    netVisual_aggregate(object.list[[i]], signaling = pathways.show, top = 0.1,layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object.list)[i]))
}


rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage"), targets.use = NULL, stacked = F, do.stat = TRUE,comparison = c(1, 2, 3),signaling=c("WNT"))

###CHECK SIG VALUE!!

a<-rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage"), targets.use = NULL, stacked = T, do.stat = TRUE,comparison = c(1, 2),return.data = TRUE)
a$signaling.contribution%>% filter(name == "WNT")









###TGFb

pathways.show <- c("TGFb") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))


netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("Tissue-resident Macrophage","Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), targets.use = c("Tissue-resident Macrophage","Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), angle.x = 45,remove.isolate = FALSE,signaling=c("TGFb"),dot.size.min =4)

netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("Tissue-resident Macrophage","Macrophage","POSTN+ FIB"), targets.use = c("Tissue-resident Macrophage","Macrophage","PRG4+ FIB","CHI3L2+ FIB","POSTN+ FIB","CXCL12+ FIB","MFAP5+ FIB"), angle.x = 45,remove.isolate = FALSE,signaling=c("TGFb"),dot.size.min =4)



plotGeneExpression(cellchat, signaling = "TGFb", split.by = "datasets", colors.ggplot = T, type = "violin")



#Circle plots
pathways.show <- c("TGFb") 
ct<-c("Tissue-resident Macrophage","Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB")
weight.max <- getMaxWeight(object.list, slot.name = c("netP"), attribute = pathways.show) # control the edge weights across different datasets
par(mfrow = c(1,1), xpd=TRUE)
for (i in 1:length(object.list)) {
    netVisual_aggregate(object.list[[i]], signaling = pathways.show,layout = "circle", edge.weight.max = weight.max[1], edge.width.max = 10, signaling.name = paste(pathways.show, names(object.list)[i]), sources.use= ct, targets.use=ct)
}


rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = ct, targets.use = ct, stacked = F, do.stat = TRUE,comparison = c(1, 2, 3),signaling=c("TGFb"))

###CHECK SIG VALUE!!

a<-rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), targets.use = NULL, stacked = T, do.stat = TRUE,comparison = c(1, 2),return.data = TRUE)
a$signaling.contribution%>% filter(name == "TGFb")

```


```{r}
#cellchat@meta$datasets = factor(cellchat@meta$datasets, levels = c("NL", "LS")) # set factor level
plotGeneExpression(cellchat, signaling = "SPP1", split.by = "datasets", colors.ggplot = T, type = "violin")

plotGeneExpression(cellchat, signaling = "IL6", split.by = "datasets", colors.ggplot = T, type = "violin")

plotGeneExpression(cellchat, signaling = "PLAU", split.by = "datasets", colors.ggplot = T, type = "violin")

plotGeneExpression(cellchat, signaling = "SELE", split.by = "datasets", colors.ggplot = T, type = "violin")

plotGeneExpression(cellchat, signaling = "COLLAGEN", split.by = "datasets", colors.ggplot = T, type = "violin")
#plotGeneExpression(cellchat, signaling = "STAT3", split.by = "datasets", colors.ggplot = T, type = "violin")
#try SMAD3
```

