---
title: "Subcluster Annotation"
subtitle: "10_CellChat"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Cell-Cell Communication

## Setup

### Load packages

```{r LoadPackages, warning=FALSE}
#Standard Packages
library(here)
source(here("code", "standard_libraries.R"))

#Additional Packages
suppressPackageStartupMessages({
library(CellChat)
library(patchwork)
})

```

### Set Parameter

```{r Setup}
set.seed(100)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=100)
path <- here::here()
```



### Load Data

```{r LoadDataset}
## RA DATASET
sce <- readRDS(file = paste0(path,'/output/07_sce_SC_Merge.rds'))

```

### Explore Dataset

```{r ShowDim}
#Dimensions of count matrix
dim(sce)
```


### Plot Annotation {.tabset}

#### Annotation L2

```{r UMAPCluster}
#cluster
plotReducedDim(sce, "MNN_merge_reduced_UMAP", colour_by="Annotation_L2") +labs(title="UMAP colored by clusters",subtitle = "UMAP of integrated dataset")

```

#### Annotation L1

```{r UMAPCluster}
#cluster
plotReducedDim(sce, "MNN_merge_reduced_UMAP", colour_by="Annotation_L2", text_by="Annotation_L2") +labs(title="UMAP colored by clusters",subtitle = "UMAP of integrated dataset")

```

#### Annotation L0

```{r UMAPAnnotation}
#cell type
plotReducedDim(sce, "MNN_merge_reduced_UMAP", colour_by="Annotation_L0", text_by="Annotation_L0") +labs(title="UMAP colored by Cell type ",subtitle = "UMAP of integrated dataset")

```


## Setup

```{r}

#Create Annotation
sce$CellChatAnno <- sce$Annotation_L2
unique(sce$CellChatAnno)

sce$CellChatAnno[sce$Annotation_L2 %in% c("CytotoxCD4 TC","Memory TC","Naive TC","CytotoxCD8 TC","NKTcell")] <- "T cell"
sce$CellChatAnno[sce$Annotation_L2 %in% c("TREM2+ MP","FOLR2+ MP")] <- "Tissue-resident Macrophage"
sce$CellChatAnno[sce$Annotation_L2 %in% c("S100A12+ MP","CD48+ MP")] <- "Macrophage"
sce$CellChatAnno[sce$Annotation_L2 %in% c("cDC","pDC")] <- "Dendritic cell"
sce$CellChatAnno[sce$Annotation_L2 %in% c("Arteriolar EC","Venular EC","Capillary EC")] <- "Endothelial"

#Create a CellChat object
cellChat.MCP <- createCellChat(object = sce[,sce$Joint.Location == "MCP"], group.by = "CellChatAnno")
cellChat.Knee <- createCellChat(object = sce[,sce$Joint.Location == "Knee"], group.by = "CellChatAnno")
cellChat.Wrist <- createCellChat(object = sce[,sce$Joint.Location == "Wrist"], group.by = "CellChatAnno")

#unique(cellchat@meta$CellChatAnno)
```

## Set the ligand-receptor interaction database

```{r}
CellChatDB <- CellChatDB.human 
showDatabaseCategory(CellChatDB)

# use all CellChatDB for cell-cell communication analysis
#CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling", key = "annotation") # use Secreted Signaling
CellChatDB.use <- subsetDB(CellChatDB) # use all Signaling


# use all CellChatDB except for "Non-protein Signaling" for cell-cell communication analysis
# CellChatDB.use <- subsetDB(CellChatDB)

# set the used database in the object
cellChat.MCP@DB <- CellChatDB.use
cellChat.Knee@DB <- CellChatDB.use
cellChat.Wrist@DB <- CellChatDB.use
```

## Preprocessing expression data for cell-cell communication analysis

```{r}
# subset the expression data of signaling genes for saving computation cost
cellChat.MCP <- subsetData(cellChat.MCP) # This step is necessary even if using the whole database
cellChat.Knee <- subsetData(cellChat.Knee)
cellChat.Wrist <- subsetData(cellChat.Wrist)

future::plan("multisession", workers = 4) # do parallel

cellChat.MCP <- identifyOverExpressedGenes(cellChat.MCP)
cellChat.Knee <- identifyOverExpressedGenes(cellChat.Knee)
cellChat.Wrist <- identifyOverExpressedGenes(cellChat.Wrist)

cellChat.MCP <- identifyOverExpressedInteractions(cellChat.MCP)
cellChat.Knee <- identifyOverExpressedInteractions(cellChat.Knee)
cellChat.Wrist <- identifyOverExpressedInteractions(cellChat.Wrist)

#cellChat.MCP <- projectData(cellChat.MCP, PPI.human)
#cellChat.Knee <- projectData(cellChat.Knee, PPI.human)
#cellChat.Wrist <- projectData(cellChat.Wrist, PPI.human)
```


## Inference of cell-cell communication network

```{r}
cellChat.MCP <- computeCommunProb(cellChat.MCP, type = "triMean")
cellChat.Knee <- computeCommunProb(cellChat.Knee, type = "triMean")
cellChat.Wrist <- computeCommunProb(cellChat.Wrist, type = "triMean")
```

```{r}
cellChat.MCP <- filterCommunication(cellChat.MCP, min.cells = 10)
cellChat.Knee <- filterCommunication(cellChat.Knee, min.cells = 10)
cellChat.Wrist <- filterCommunication(cellChat.Wrist, min.cells = 10)
```

## Infer the cell-cell communication at a signaling pathway level

```{r}
cellChat.MCP <- computeCommunProbPathway(cellChat.MCP)
cellChat.Knee <- computeCommunProbPathway(cellChat.Knee)
cellChat.Wrist <- computeCommunProbPathway(cellChat.Wrist)

cellChat.MCP <- aggregateNet(cellChat.MCP)
cellChat.Knee <- aggregateNet(cellChat.Knee)
cellChat.Wrist <- aggregateNet(cellChat.Wrist)

cellChat.MCP<-netAnalysis_computeCentrality(cellChat.MCP)
cellChat.Knee<-netAnalysis_computeCentrality(cellChat.Knee)
cellChat.Wrist<-netAnalysis_computeCentrality(cellChat.Wrist)
```

## Save the datasets

```{r SaveData}
saveRDS(cellChat.MCP, file =paste0(path,'/output/10_sce_CellChat_MCP.rds'))
saveRDS(cellChat.Knee, file =paste0(path,'/output/10_sce_CellChat_Knee.rds'))
saveRDS(cellChat.Wrist, file =paste0(path,'/output/10_sce_CellChat_Wrist.rds'))
```



## Load CellChat object of each dataset and merge them together

### Load CellChat
```{r}
cellChat.MCP <- readRDS(file =paste0(path,'/output/10_sce_CellChat_MCP.rds'))
cellChat.Knee <- readRDS(file =paste0(path,'/output/10_sce_CellChat_Knee.rds'))
cellChat.Wrist <- readRDS(file =paste0(path,'/output/10_sce_CellChat_Wrist.rds'))
```


### Merge CellChat objects

```{r}
cellChat.MCP<-updateCellChat(cellChat.MCP)
cellChat.Knee<-updateCellChat(cellChat.Knee)
cellChat.Wrist<-updateCellChat(cellChat.Wrist)

object.list <- list(MCP = cellChat.MCP, Knee = cellChat.Knee, Wrist = cellChat.Wrist)#list(MCP = cellChat.MCP, Knee = cellChat.Knee, Wrist = cellChat.Wrist)
cellchat <- mergeCellChat(object.list, add.names = names(object.list),cell.prefix = TRUE)
cellchat
```


## Comparison Analysis

### Compare the total number of interactions and interaction strength
```{r}
gg1 <- compareInteractions(cellchat, show.legend = F, group = c("MCP","Knee","Wrist"))
gg2 <- compareInteractions(cellchat, show.legend = F, group = c("MCP","Knee","Wrist"), measure = "weight")
gg1 + gg2
```


```{r}
# Extract differential interactions
interaction_diff <- cellchat@net$compare$result
head(interaction_diff)
```


### Differential number of interactions or interaction strength among coarse cell types
```{r}
#levels(cellchat@idents)
#group.cellType <- c(rep("FIB", 4), rep("DC", 4), rep("TC", 4))
#group.cellType <- factor(group.cellType, levels = c("FIB", "DC", "TC"))
#object.list <- lapply(object.list, function(x) {mergeInteractions(x, group.cellType)})
#cellchat <- mergeCellChat(object.list, add.names = names(object.list))



#par(mfrow = c(1,2), xpd=TRUE)
#for (i in 1:length(object.list)) {
  #netVisual_circle(object.list[[i]]@net$count.merged, weight.scale = T, label.edge= T, edge.weight.max = weight.max[3], edge.width.max = 12, title.name = paste0("Number of interactions - ", names(object.list)[i]))
#}
```




### Compare the number of interactions and interaction strength among different cell populations
```{r}
#sources.use =c("CHI3L2+ FIB","CXCL12+ FIB","MFAP5+ FIB","POSTN+ FIB","PRG4+ FIB")
#targets.use = c("FOLR2+ MP","TREM2+ MP")

#KneeVsMCP
gg1 <- netVisual_heatmap(cellchat,comparison = c("Knee","MCP"),title.name = "Knee Vs MCP")
gg2 <- netVisual_heatmap(cellchat,comparison = c("Knee","MCP"), measure = "weight",title.name = "Knee Vs MCP")
gg1 + gg2

#KneeVsWrist
gg1 <- netVisual_heatmap(cellchat,comparison = c("Knee","Wrist"),title.name = "Knee Vs Wrist")
gg2 <- netVisual_heatmap(cellchat,comparison = c("Knee","Wrist"), measure = "weight",title.name = "Knee Vs Wrist")
gg1 + gg2

#WRISTVsMCP
gg1 <- netVisual_heatmap(cellchat,comparison = c("MCP","Wrist"),title.name = "MCP Vs Wrist")
gg2 <- netVisual_heatmap(cellchat, comparison = c("MCP","Wrist"), measure = "weight",title.name = "MCP Vs Wrist")
gg1 + gg2
```


### Identify cell populations with significant changes in sending or receiving signals

```{r}

num.link <- sapply(object.list, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg <- list()
for (i in 1:length(object.list)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(object.list[[i]], title = names(object.list)[i], weight.MinMax = weight.MinMax)
}
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
patchwork::wrap_plots(plots = gg)
```

```{r}
#POSTN+ FIB
gg1 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 1), idents.use = "PRG4+ FIB", signaling.exclude = c("MIF"))
gg2 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 3), idents.use = "PRG4+ FIB", signaling.exclude = c("MIF"))
gg3 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(1, 3), idents.use = "PRG4+ FIB", signaling.exclude = c("MIF"))


patchwork::wrap_plots(plots = list(gg1,gg2,gg3))


#Tissue-resident Macrophage 
gg1 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 1), idents.use = "Tissue-resident Macrophage", signaling.exclude = c("MIF"))
gg2 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 3), idents.use = "Tissue-resident Macrophage", signaling.exclude = c("MIF"))
gg3 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(1, 3), idents.use = "Tissue-resident Macrophage", signaling.exclude = c("MIF"))


patchwork::wrap_plots(plots = list(gg1,gg2,gg3))


#CHI3L2+ FIB
gg1 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 1), idents.use = "CHI3L2+ FIB", signaling.exclude = c("MIF"))
gg2 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 3), idents.use = "CHI3L2+ FIB", signaling.exclude = c("MIF"))
gg3 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(1, 3), idents.use = "CHI3L2+ FIB", signaling.exclude = c("MIF"))


#patchwork::wrap_plots(plots = list(gg1,gg2,gg3))


#CHI3L2+ FIB
gg1 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 1), idents.use = "Neutrophil", signaling.exclude = c("MIF"))
gg2 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(2, 3), idents.use = "Neutrophil", signaling.exclude = c("MIF"))
gg3 <- netAnalysis_signalingChanges_scatter(cellchat, comparison = c(1, 3), idents.use = "Neutrophil", signaling.exclude = c("MIF"))


patchwork::wrap_plots(plots = list(gg1,gg2,gg3))

```

### Identify cell populations with significant changes in sending or receiving signals

```{r}
pathways.show <- c("SEMA3") 
par(mfrow = c(1,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
  ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))
```


### Identify cell populations with significant changes in sending or receiving signals

```{r}
gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE,comparison = c(1, 2))
gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE,comparison = c(1, 2))

gg1 + gg2


gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE,comparison = c(1, 3))
gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE,comparison = c(1, 3))

gg1 + gg2


gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE,comparison = c(3, 2))
gg2 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE,comparison = c(3, 2))

gg1 + gg2

rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = F, do.stat = TRUE,comparison = c(1, 2, 3))

rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial"), targets.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial"), stacked = F, do.stat = TRUE,comparison = c(1, 2, 3))


rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial"), targets.use = c("Tissue-resident Macrophage","POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Smooth muscle cell","Endothelial"), stacked = F, do.stat = TRUE,comparison = c(1, 2, 3),signaling=c("SEMA3","GAP","ncWNT","WNT","MMP","PTN","CXCL","FGF","EGF","GAP","ANGPTL","SELE","NOTCH","VEGEF","TGFB","IL16","ADGRL","PDGF","MIF"))

```


```{r}
#cellchat@meta$datasets = factor(cellchat@meta$datasets, levels = c("NL", "LS")) # set factor level
netVisual_bubble(cellchat, sources.use = c("Tissue-resident Macrophage"), targets.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB"), comparison = c(1, 2), angle.x = 45,remove.isolate = FALSE)

netVisual_bubble(cellchat, comparison = c(1, 2),sources.use = c("CXCL12+ FIB","PRG4+ FIB","Endothelial","Tissue-resident Macrophage"), targets.use = c("CXCL12+ FIB","PRG4+ FIB","Endothelial","Tissue-resident Macrophage",), angle.x = 45,remove.isolate = FALSE,signaling=c("CXCL","IL6"))

netVisual_bubble(cellchat, comparison = c(1, 2, 3),sources.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Endothelial","Tissue-resident Macrophage","Neutrophil"), targets.use = c("POSTN+ FIB","CXCL12+ FIB","PRG4+ FIB","CHI3L2+ FIB","MFAP5+ FIB","Endothelial","Tissue-resident Macrophage","Neutrophil"), angle.x = 45,remove.isolate = FALSE,signaling=c("ICAM"))

#try SMAD3





pathways.show <- c("CXCL") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 

#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))



pathways.show <- c("IL6") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))




pathways.show <- c("WNT") 
par(mfrow = c(2,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
    ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = pathways.show, color.heatmap = "Reds",title.name = paste(pathways.show, "signaling ",names(object.list)[i]))
}
#> Do heatmap based on a single object 
#> 
#> Do heatmap based on a single object
ComplexHeatmap::draw(ht[[1]] + ht[[2]]+ ht[[3]], ht_gap = unit(0.5, "cm"))

```


```{r}
#cellchat@meta$datasets = factor(cellchat@meta$datasets, levels = c("NL", "LS")) # set factor level
plotGeneExpression(cellchat, signaling = "SPP1", split.by = "datasets", colors.ggplot = T, type = "violin")

plotGeneExpression(cellchat, signaling = "IL6", split.by = "datasets", colors.ggplot = T, type = "violin")

plotGeneExpression(cellchat, signaling = "PLAU", split.by = "datasets", colors.ggplot = T, type = "violin")

plotGeneExpression(cellchat, signaling = "SELE", split.by = "datasets", colors.ggplot = T, type = "violin")

plotGeneExpression(cellchat, signaling = "COLLAGEN", split.by = "datasets", colors.ggplot = T, type = "violin")
#plotGeneExpression(cellchat, signaling = "STAT3", split.by = "datasets", colors.ggplot = T, type = "violin")
#try SMAD3
```





```{r}
library(Seurat)
library(scDiffCom)
#library(data.table)

sce.sc<-sce[,sce$Joint.Location != "Wrist"]

so <- as.Seurat(sce.sc, counts = "counts", data = "logcounts" )
#so[['RNA']] = so[['originalexp']]
so

scdiffcom_object <- run_interaction_analysis(
  seurat_object = so,
  LRI_species = "human",
  seurat_assay = "originalexp",
  seurat_slot = "data",
  seurat_celltype_id = "CellChatAnno",
  seurat_condition_id = list(
    column_name = "Joint.Location",
    cond1_name = "MCP",
    cond2_name = "Knee"
  )
)

```

```{r}

scdiffcom_object

CCI_detected <- GetTableCCI(scdiffcom_object, type = "detected", simplified = TRUE)
table(CCI_detected$REGULATION)

ORA_results <- GetTableORA(scdiffcom_object, categories = "all", simplified = TRUE)
names(ORA_results)


ggplot(CCI_detected,aes(x = LOGFC,y = -log10(BH_P_VALUE_DE + 1E-2),colour = REGULATION)) + 
  geom_point() + 
  scale_colour_manual(values = c("UP" = "red", "DOWN" = "blue", "FLAT" = "green", "NSC" = "grey"))+ 
  xlab("log(FC)") + 
  ylab("-log10(Adj. p-value)")

PlotORA(object = scdiffcom_object,category = "LRI",regulation = "UP") + 
  theme(legend.position = c(0.85, 0.4),legend.key.size = unit(0.4, "cm"))


```




```{r}

sce.sc<-sce[,sce$Joint.Location != "MCP"]

so <- as.Seurat(sce.sc, counts = "counts", data = "logcounts" )
#so[['RNA']] = so[['originalexp']]
so

scdiffcom_object <- run_interaction_analysis(
  seurat_object = so,
  LRI_species = "human",
  seurat_assay = "originalexp",
  seurat_slot = "data",
  seurat_celltype_id = "CellChatAnno",
  seurat_condition_id = list(
    column_name = "Joint.Location",
    cond1_name = "Wrist",
    cond2_name = "Knee"
  )
)

```


```{r}

scdiffcom_object

CCI_detected <- GetTableCCI(scdiffcom_object, type = "detected", simplified = TRUE)
table(CCI_detected$REGULATION)

ORA_results <- GetTableORA(scdiffcom_object, categories = "all", simplified = TRUE)
names(ORA_results)


ggplot(CCI_detected,aes(x = LOGFC,y = -log10(BH_P_VALUE_DE + 1E-2),colour = REGULATION)) + 
  geom_point() + 
  scale_colour_manual(values = c("UP" = "red", "DOWN" = "blue", "FLAT" = "green", "NSC" = "grey"))+ 
  xlab("log(FC)") + 
  ylab("-log10(Adj. p-value)")

PlotORA(object = scdiffcom_object,category = "LRI",regulation = "UP") + 
  theme(legend.position = c(0.85, 0.4),legend.key.size = unit(0.4, "cm"))


```



```{r}

sce.sc<-sce[,sce$Joint.Location != "Knee"]

so <- as.Seurat(sce.sc, counts = "counts", data = "logcounts" )
#so[['RNA']] = so[['originalexp']]
so

scdiffcom_object <- run_interaction_analysis(
  seurat_object = so,
  LRI_species = "human",
  seurat_assay = "originalexp",
  seurat_slot = "data",
  seurat_celltype_id = "CellChatAnno",
  seurat_condition_id = list(
    column_name = "Joint.Location",
    cond1_name = "MCP",
    cond2_name = "Wrist"
  )
)

```


```{r}

scdiffcom_object

CCI_detected <- GetTableCCI(scdiffcom_object, type = "detected", simplified = TRUE)
table(CCI_detected$REGULATION)

ORA_results <- GetTableORA(scdiffcom_object, categories = "all", simplified = TRUE)
names(ORA_results)


ggplot(CCI_detected,aes(x = LOGFC,y = -log10(BH_P_VALUE_DE + 1E-2),colour = REGULATION)) + 
  geom_point() + 
  scale_colour_manual(values = c("UP" = "red", "DOWN" = "blue", "FLAT" = "green", "NSC" = "grey"))+ 
  xlab("log(FC)") + 
  ylab("-log10(Adj. p-value)")

PlotORA(object = scdiffcom_object,category = "LRI",regulation = "UP") + 
  theme(legend.position = c(0.85, 0.4),legend.key.size = unit(0.4, "cm"))


```


#---------


```{r}
library(nichenetr)
library(multinichenetr)

# Load NicheNet’s ligand-receptor network and ligand-target matrix

organism = "human"

lr_network_all = readRDS(url("https://zenodo.org/record/10229222/files/lr_network_human_allInfo_30112033.rds")) %>% 
    mutate(ligand = convert_alias_to_symbols(ligand, organism = organism), receptor = convert_alias_to_symbols(receptor, organism = organism))
  
lr_network_all = lr_network_all  %>% 
  mutate(ligand = make.names(ligand), receptor = make.names(receptor)) 
  
lr_network = lr_network_all %>% 
  distinct(ligand, receptor)
  
ligand_target_matrix = readRDS(url("https://zenodo.org/record/7074291/files/ligand_target_matrix_nsga2r_final.rds"))
  
colnames(ligand_target_matrix) = colnames(ligand_target_matrix) %>% 
  convert_alias_to_symbols(organism = organism) %>% make.names()
rownames(ligand_target_matrix) = rownames(ligand_target_matrix) %>% 
  convert_alias_to_symbols(organism = organism) %>% make.names()
  
lr_network = lr_network %>% filter(ligand %in% colnames(ligand_target_matrix))
ligand_target_matrix = ligand_target_matrix[, lr_network$ligand %>% unique()]


#Create Annotation
sce$CellChatAnno <- sce$Annotation_L2
unique(sce$CellChatAnno)
sce$CellChatAnno[sce$Annotation_L2 %in% c("CytotoxCD4 TC","Memory TC","Naive TC","CytotoxCD8 TC","NKTcell")] <- "T cell"
sce$CellChatAnno[sce$Annotation_L2 %in% c("TREM2+ MP","FOLR2+ MP")] <- "Tissueresident Macrophage"
sce$CellChatAnno[sce$Annotation_L2 %in% c("S100A12+ MP","CD48+ MP")] <- "Macrophage"
sce$CellChatAnno[sce$Annotation_L2 %in% c("cDC","pDC")] <- "Dendritic cell"
sce$CellChatAnno[sce$Annotation_L2 %in% c("Arteriolar EC","Venular EC","Capillary EC")] <- "Endothelial"


# Update genesymbols
sce = alias_to_symbol_SCE(sce, "human") %>% makenames_SCE()

sample_id = "Sample"
group_id = "Joint.Location"
celltype_id = "CellChatAnno"

#Make names compatible
colData(sce)[,celltype_id]<-make.names(colData(sce)[,celltype_id])



#Set contrast
#contrasts_oi = c("'Knee-MCP','Knee-Wrist','MCP-Wrist'")
contrasts_oi = c("'Knee-(MCP+Wrist)/2','MCP-(Knee+Wrist)/2','Wrist-(Knee+MCP)/2'")

contrast_tbl = tibble(contrast = 
                        c("Knee-(MCP+Wrist)/2","MCP-(Knee+Wrist)/2", "Wrist-(Knee+MCP)/2"), 
                      group = c("Knee","MCP","Wrist"))


#Define the sender and receiver cell types of interest
senders_oi = colData(sce)[,celltype_id] %>% unique()
receivers_oi = colData(sce)[,celltype_id] %>% unique()
sce = sce[, colData(sce)[,celltype_id] %in% c(senders_oi, receivers_oi)]

#Cell-type filtering
min_cells = 5
abundance_info = get_abundance_info(
  sce = sce, 
  sample_id = sample_id, group_id = group_id, celltype_id = celltype_id, 
  min_cells = min_cells, 
  senders_oi = senders_oi, receivers_oi = receivers_oi, 
  #batches = batches
  )

abundance_info$abund_plot_sample

```


```{r}
sample_group_celltype_df = abundance_info$abundance_data %>% 
  filter(n > min_cells) %>% 
  ungroup() %>% 
  distinct(sample_id, group_id) %>% 
  cross_join(
    abundance_info$abundance_data %>% 
      ungroup() %>% 
      distinct(celltype_id)
    ) %>% 
  arrange(sample_id)

abundance_df = sample_group_celltype_df %>% left_join(
  abundance_info$abundance_data %>% ungroup()
  )

abundance_df$n[is.na(abundance_df$n)] = 0
abundance_df$keep[is.na(abundance_df$keep)] = FALSE
abundance_df_summarized = abundance_df %>% 
  mutate(keep = as.logical(keep)) %>% 
  group_by(group_id, celltype_id) %>% 
  summarise(samples_present = sum((keep)))

celltypes_absent_one_condition = abundance_df_summarized %>% 
  filter(samples_present == 0) %>% pull(celltype_id) %>% unique() 
# find truly condition-specific cell types by searching for cell types 
# truely absent in at least one condition

celltypes_present_one_condition = abundance_df_summarized %>% 
  filter(samples_present >= 2) %>% pull(celltype_id) %>% unique() 
# require presence in at least 2 samples of one group so 
# it is really present in at least one condition

condition_specific_celltypes = intersect(
  celltypes_absent_one_condition, 
  celltypes_present_one_condition)

total_nr_conditions = SummarizedExperiment::colData(sce)[,group_id] %>% 
  unique() %>% length() 

absent_celltypes = abundance_df_summarized #%>% 
  #filter(samples_present < 2) %>% 
  #group_by(celltype_id) %>% 
  #count() %>% 
  #filter(n == total_nr_conditions) %>% 
  #pull(celltype_id)
  
print("condition-specific celltypes:")
print(condition_specific_celltypes)

  
print("absent celltypes:")
print(absent_celltypes)

```


```{r}
#Gene filtering
min_sample_prop = 0.50
fraction_cutoff = 0.05

frq_list = get_frac_exprs(
  sce = sce, 
  sample_id = sample_id, celltype_id =  celltype_id, group_id = group_id, 
  #batches = batches, 
  min_cells = min_cells, 
  fraction_cutoff = fraction_cutoff, min_sample_prop = min_sample_prop)

genes_oi = frq_list$expressed_df %>% 
  filter(expressed == TRUE) %>% pull(gene) %>% unique() 
sce = sce[genes_oi, ]


#Pseudobulk expression calculation
abundance_expression_info = process_abundance_expression_info(
  sce = sce, 
  sample_id = sample_id, group_id = group_id, celltype_id = celltype_id, 
  min_cells = min_cells, 
  senders_oi = senders_oi, receivers_oi = receivers_oi, 
  lr_network = lr_network, 
  #batches = batches, 
  frq_list = frq_list, 
  abundance_info = abundance_info)

#Inspect
abundance_expression_info$celltype_info$pb_df %>% head()
abundance_expression_info$celltype_info$pb_df_group %>% head()
abundance_expression_info$sender_receiver_info$pb_df %>% head()
abundance_expression_info$sender_receiver_info$pb_df_group %>% head()

```


```{r}
#Differential expression (DE) analysis
DE_info = get_DE_info(
  sce = sce, 
  sample_id = sample_id, group_id = group_id, celltype_id = celltype_id, 
  batches = NA, covariates = NA, 
  contrasts_oi = contrasts_oi, 
  min_cells = min_cells, 
  expressed_df = frq_list$expressed_df)
#check results
DE_info$celltype_de$de_output_tidy %>% head()
DE_info$hist_pvals

DE_info_emp = get_empirical_pvals(DE_info$celltype_de$de_output_tidy)
celltype_de = DE_info_emp$de_output_tidy_emp %>% dplyr::select(-p_val, -p_adj) %>% 
  rename(p_val = p_emp, p_adj = p_adj_emp)

#Combine DE information for ligand-senders and receptors-receivers

sender_receiver_de = combine_sender_receiver_de(
  sender_de = celltype_de,
  receiver_de = celltype_de,
  senders_oi = senders_oi,
  receivers_oi = receivers_oi,
  lr_network = lr_network
)

```

```{r}
#Ligand activity prediction

logFC_threshold = 0.50
p_val_threshold = 0.05

p_val_adj = FALSE 

geneset_assessment = contrast_tbl$contrast %>% 
  lapply(process_geneset_data, celltype_de, logFC_threshold, p_val_adj, p_val_threshold) %>% 
  bind_rows() 
geneset_assessment

#adjustedPval
geneset_assessment_adjustedPval = contrast_tbl$contrast %>% 
  lapply(
    process_geneset_data, 
    celltype_de, logFC_threshold, p_val_adj = TRUE, p_val_threshold
    ) %>% 
  bind_rows() 
geneset_assessment_adjustedPval


# Perform the ligand activity analysis and ligand-target inference
top_n_target = 250
verbose = TRUE
cores_system = 8
n.cores = min(cores_system, celltype_de$cluster_id %>% unique() %>% length()) 

ligand_activities_targets_DEgenes = suppressMessages(suppressWarnings(
  get_ligand_activities_targets_DEgenes(
    receiver_de = celltype_de,
    receivers_oi = intersect(receivers_oi, celltype_de$cluster_id %>% unique()),
    ligand_target_matrix = ligand_target_matrix,
    logFC_threshold = logFC_threshold,
    p_val_threshold = p_val_threshold,
    p_val_adj = p_val_adj,
    top_n_target = top_n_target,
    verbose = verbose, 
    n.cores = n.cores
  )
))
ligand_activities_targets_DEgenes$ligand_activities %>% head(20)


```


```{r}
ligand_activity_down = FALSE
sender_receiver_tbl = sender_receiver_de %>% distinct(sender, receiver)

metadata_combined = SummarizedExperiment::colData(sce) %>% tibble::as_tibble()


grouping_tbl = metadata_combined[,c(sample_id, group_id)] %>% 
tibble::as_tibble() %>% distinct()
colnames(grouping_tbl) = c("sample","group")


prioritization_tables = suppressMessages(generate_prioritization_tables(
    sender_receiver_info = abundance_expression_info$sender_receiver_info,
    sender_receiver_de = sender_receiver_de,
    ligand_activities_targets_DEgenes = ligand_activities_targets_DEgenes,
    contrast_tbl = contrast_tbl,
    sender_receiver_tbl = sender_receiver_tbl,
    grouping_tbl = grouping_tbl,
    scenario = "regular", # all prioritization criteria will be weighted equally
    fraction_cutoff = fraction_cutoff, 
    abundance_data_receiver = abundance_expression_info$abundance_data_receiver,
    abundance_data_sender = abundance_expression_info$abundance_data_sender,
    ligand_activity_down = ligand_activity_down
  ))

prioritization_tables$group_prioritization_tbl %>% head(20)

#Calculate the across-samples expression correlation between ligand-receptor pairs and target genes
lr_target_prior_cor = lr_target_prior_cor_inference(
  receivers_oi = prioritization_tables$group_prioritization_tbl$receiver %>% unique(), 
  abundance_expression_info = abundance_expression_info, 
  celltype_de = celltype_de, 
  grouping_tbl = grouping_tbl, 
  prioritization_tables = prioritization_tables, 
  ligand_target_matrix = ligand_target_matrix, 
  logFC_threshold = logFC_threshold, 
  p_val_threshold = p_val_threshold, 
  p_val_adj = p_val_adj
  )


multinichenet_output = list(
    celltype_info = abundance_expression_info$celltype_info,
    celltype_de = celltype_de,
    sender_receiver_info = abundance_expression_info$sender_receiver_info,
    sender_receiver_de =  sender_receiver_de,
    ligand_activities_targets_DEgenes = ligand_activities_targets_DEgenes,
    prioritization_tables = prioritization_tables,
    grouping_tbl = grouping_tbl,
    lr_target_prior_cor = lr_target_prior_cor
  ) 
multinichenet_output = make_lite_output(multinichenet_output)


saveRDS(multinichenet_output, paste0(path, "/output/","multinichenet_output.rds"))


```


```{r}
#Plots
send<-c("Smooth.muscle.cell","Neutrophil","POSTN..FIB","CXCL12..FIB","PRG4..FIB","CHI3L2..FIB","MFAP5..FIB","Tissueresident.Macrophage","T.cell","Endothelial")
rec<-c("Smooth.muscle.cell","Neutrophil","POSTN..FIB","CXCL12..FIB","PRG4..FIB","CHI3L2..FIB","MFAP5..FIB","Tissueresident.Macrophage","T.cell","Endothelial")

rec<-c("Endothelial")
send<-c("Smooth.muscle.cell","Neutrophil","POSTN..FIB","CXCL12..FIB","PRG4..FIB","CHI3L2..FIB","MFAP5..FIB","Tissueresident.Macrophage","T.cell","Endothelial")

prioritized_tbl_oi_all = get_top_n_lr_pairs(
  multinichenet_output$prioritization_tables, 
  top_n = 50, 
  senders_oi = send, 
  receivers_oi = rec,
  rank_per_group = FALSE
  )

#ChordDiagram
prioritized_tbl_oi = 
  multinichenet_output$prioritization_tables$group_prioritization_tbl %>%
  filter(id %in% prioritized_tbl_oi_all$id) %>%
  distinct(id, sender, receiver, ligand, receptor, group) %>% 
  left_join(prioritized_tbl_oi_all)
prioritized_tbl_oi$prioritization_score[is.na(prioritized_tbl_oi$prioritization_score)] = 0

senders_receivers = union(prioritized_tbl_oi$sender %>% unique(), prioritized_tbl_oi$receiver %>% unique()) %>% sort()

colors_sender = RColorBrewer::brewer.pal(n = length(senders_receivers), name = 'Spectral') %>% magrittr::set_names(senders_receivers)
colors_receiver = RColorBrewer::brewer.pal(n = length(senders_receivers), name = 'Spectral') %>% magrittr::set_names(senders_receivers)

circos_list = make_circos_group_comparison(prioritized_tbl_oi, colors_sender, colors_receiver)


#bubble plots
group_oi = "MCP"
prioritized_tbl_oi_M_50 = prioritized_tbl_oi_all %>% 
  filter(group == group_oi)
plot_oi = make_sample_lr_prod_activity_plots_Omnipath(
  multinichenet_output$prioritization_tables, 
  prioritized_tbl_oi_M_50 %>% inner_join(lr_network_all)
  )
plot_oi


```


```{r}
#Plots
send<-c("Smooth.muscle.cell","Neutrophil","POSTN..FIB","CXCL12..FIB","PRG4..FIB","CHI3L2..FIB","MFAP5..FIB","Tissueresident.Macrophage","T.cell","Endothelial")
rec<-c("Smooth.muscle.cell","Neutrophil","POSTN..FIB","CXCL12..FIB","PRG4..FIB","CHI3L2..FIB","MFAP5..FIB","Tissueresident.Macrophage","T.cell","Endothelial")

rec<-c("Endothelial")
send<-c("Smooth.muscle.cell","Neutrophil","POSTN..FIB","CXCL12..FIB","PRG4..FIB","CHI3L2..FIB","MFAP5..FIB","Tissueresident.Macrophage","T.cell","Endothelial")

prioritized_tbl_oi_all = get_top_n_lr_pairs(
  multinichenet_output$prioritization_tables, 
  top_n = 100, 
  senders_oi = send, 
  receivers_oi = rec,
  rank_per_group = FALSE
  )


#
lr_target_prior = prioritized_tbl_oi_all %>% inner_join(
        multinichenet_output$ligand_activities_targets_DEgenes$ligand_activities %>%
          distinct(ligand, target, direction_regulation, contrast) %>% inner_join(contrast_tbl) %>% ungroup() 
        ) 
lr_target_df = lr_target_prior %>% distinct(group, sender, receiver, ligand, receptor, id, target, direction_regulation) 

lr_target_df %>% filter(target %in% union(lr_network$ligand, lr_network$receptor))

network = infer_intercellular_regulatory_network(lr_target_df, prioritized_tbl_oi_all)
network$links %>% head()


network$nodes %>% head()

#colors_sender["L_T_TIM3._CD38._HLADR."] = "pink" # the  original yellow background with white font is not very readable
network_graph = visualize_network(network, colors_sender)
network_graph$plot

network$nodes %>% filter(gene == "SELE")

```
