---
title: "Spatial Analysis Knee"
subtitle: "Spatial_knee"
author: "SarahL."
date: "`r format(Sys.time(), '%b %e %Y %H:%M:%S')`"
output: 
  html_document:
    toc: true
    number_sections: true   
    toc_float: true
    code_folding: hide
    df_print: paged
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(cache = FALSE, cache.lazy = FALSE)
knitr::opts_chunk$set(dpi = 300, fig.align = "center")
knitr::opts_chunk$set(out.width = "100%", out.height = "100%")
```


# Spatial Analysis Knee

## Setup

### Standard packages

```{r LoadPackages, warning=FALSE}
library(here)
source(here("code", "standard_libraries.R"))
```

### Additional Packages

```{r LoadPackages, warning=FALSE}
#Load Packages

suppressPackageStartupMessages({
library(ggspavis)
library(bluster)
library(SPOTlight)
library(SpatialExperiment)
library(ggcorrplot)
library(NMF)
})

```

### Set Parameter

```{r Setup}
set.seed(100)
bpp <- BiocParallel::MulticoreParam(parallel::detectCores()-1, RNGseed=100)
path <- here::here()
```


### Load Data

```{r }
spe <- read10xVisium(paste0(path,'/data/spatial/A2_SHK166'), c("A2_SHK166"), type = "sparse", data = "filtered", images = "lowres", load = FALSE)

#Edit Rownames
rowData(spe)$ENSEMBL <- rownames(spe)
rownames(spe) <- paste0(rowData(spe)$ENSEMBL, ".", rowData(spe)$symbol)
spe
```

```{r }
dim(spe)
```

```{r }
#Inspect spe object
head(colData(spe))
head(rowData(spe))
head(spatialCoords(spe))
imgData(spe)
```

```{r }
spi <- getImg(spe)

#Plot H&E image
plot(imgRaster(spi))

#Plot spatial coordinates (spots)
plotSpots(spe)
```
## Pre-processing


```{r }
# subset to keep only spots over tissue
spe <- spe[, colData(spe)$in_tissue == 1]
dim(spe)
```

 

```{r }
# Indentify mitochondrial genes
is_mito<- grepl("(^MT-)|(^mt-)", rowData(spe)$symbol)
cat("Mitochondrial genes:", rowData(spe)$symbol[is_mito]) #Show the MT genes
table(is_mito)
```


```{r }
# calculate per-spot QC metrics and store in colData
spe <- addPerCellQC(spe, subsets = list(mito = is_mito))
head(colData(spe))
```



```{r }
#Plot Library size
hist(spe$sum,breaks = 100,main ="UMIs per spot",xlab ="UMI")
abline(v = 750, col = "red")
plotSpots(spe, annotate = "sum",size = 1.7)+  scale_colour_gradient(low = "lightgrey", high = "blue")+labs(title="Libsize")
#Plot Genes detected
hist(spe$detected,breaks = 100,main ="Genes per spot",xlab ="UMI")
abline(v = 500, col = "red")
plotSpots(spe, annotate = "detected",size = 1.7)+  scale_colour_gradient(low = "lightgrey", high = "blue")+labs(title="Libsize")

#Plot Genes detected
#hist(spe$subsets_Mito_percent,breaks = 100,main ="Genes per spot",xlab ="UMI")
#abline(v = 400, col = "red")
#plotVisium(spe, fill = "subsets_Mito_percent", highlight = "in_tissue",image = FALSE, pal = "darkred") + guides(fill = guide_colorbar(title = "Libsize"))

#Plot library size vs. number of cells per spot
#plotSpots(spe, type = "scatter",metric_x = "sum")

```

```{r }
# select QC thresholds
colData(spe)$qc.count_manual <- spe$sum < 200  
colData(spe)$qc.genexp_manual <- spe$detected < 200 
#qc.mito_manual <- sce$subsets_Mito_percent > 15 #Mitochondrial contamination has to be lower than 15%


# combined set of discarded spots
colData(spe)$discard <- spe$qc.count_manual | spe$qc.genexp_manual #qc_mito 
table(colData(spe)$discard)

# QCplots
print(plotQC(spe, type = "spot", discard = "qc.count_manual"))
print(plotQC(spe, type = "spot", discard = "qc.genexp_manual"))
print(plotQC(spe, type = "spot", discard = "discard"))
```


```{r }
# filter low-quality spots
spe <- spe[, !colData(spe)$discard]
plotSpots(spe,size = 1)
```
```{r }
# filter any new zeros created after filtering low-expressed genes

# remove genes with zero expression
spe <- spe[!rowSums(counts(spe)) == 0, ]
dim(spe)

# remove spots with zero expression
spe <- spe[, !colSums(counts(spe)) == 0]
dim(spe)
```


## Normalisation

```{r }
# normalise
spe <- computeLibraryFactors(spe)

ggplot(data.frame("log10libSf"=log10(sizeFactors(spe))), aes(x=log10libSf)) +
  xlab("Log10 Library size factor") +
  geom_histogram(bins=50)+
  ggtitle("Distribution of Library size factors")+
  theme_bw()+
  theme(panel.grid.major.y = element_blank(),panel.grid.minor.y = element_blank())


summary(sizeFactors(spe))
```
```{r }
# normalise
spe <- logNormCounts(spe)
assayNames(spe)
```


```{r }
dim(counts(spe))
```


#DimRed

```{r }
# fit mean-variance relationship
#genes <- !grepl(pattern = "^RP[L|S]|MT", x = rownames(sce))
dec <- modelGeneVar(spe, density.weights=FALSE)

top.hvgs <- getTopHVGs(dec, n=2000)

rowData(spe)$is_hvg <- rownames(spe) %in% top.hvgs

#Plot
fit <- metadata(dec)
plot(fit$mean, fit$var, col= c("#F5953C","#7A95BD")[(names(fit$mean) %in% top.hvgs)+1], xlab="Mean log-expression", 
        ylab="Variance")
    curve(fit$trend(x), col="black", add=TRUE)
    legend("topright", legend=c("Not HVG", "HVG"), col = c("#F5953C","#7A95BD"),bty="n",pch=10)
```

```{r }
# PCA
spe <- runPCA(spe, subset_row = top.hvgs)
plotReducedDim(spe, dimred = "PCA",point_alpha=1,point_size=0.5)


pca_pct_variance <- data.frame(variance = attr(reducedDim(spe, "PCA"), "percentVar"))
pca_pct_variance$PC <- 1:nrow(pca_pct_variance)

ggplot(pca_pct_variance,aes(PC, variance)) +geom_line(color="grey") +geom_point() +labs(y = "Variance explained (%)")+ geom_vline(xintercept = 15,color="red")+theme_classic()+ ggtitle("Percentage variance explained by PCs")

reducedDim(spe,"PCA") <- reducedDim(spe,"PCA")[,seq_len(15)]
```

```{r }
# UMAP
spe <- runUMAP(spe, dimred = "PCA")
plotReducedDim(spe, dimred = "UMAP",point_alpha=1,point_size=0.5)
```


## Clustering with HVGs

```{r }
colData(spe)$leiden <- clusterCells(spe, use.dimred = "PCA", BLUSPARAM = SNNGraphParam(k=25,cluster.fun="leiden",BPPARAM =bpp))
data.frame(as.list(table(colData(spe)$leiden)))

```


```{r }
plotReducedDim(spe, dimred = "UMAP",colour_by ="leiden",point_alpha=1,point_size=0.5)

# plot clusters in spatial x-y coordinates
plotSpots(spe, annotate = "leiden",size = 1)
plotVisium(spe, fill = "leiden")
```

## Deconvolution Annotation_L1

```{r }
sce <- readRDS(file = paste0(path,'/output/07_sce_SC_Merge.rds'))
#exclude some celltypes
#sce <- sce[,sce$celltype != "Mast cell"]
#sce <- sce[,sce$celltype != "Smooth muscle cell"]

```


```{r }
mgs <- scoreMarkers(sce, subset.row = rownames(sce)[rowData(sce)$hvg])
mgs_fil <- lapply(names(mgs), function(i) {
    x <- mgs[[i]]
    # Filter and keep relevant marker genes, those with AUC > 0.8
    x <- x[x$mean.AUC > 0.5, ]
    # Sort the genes from highest to lowest weight
    x <- x[order(x$mean.AUC, decreasing = TRUE), ]
    # Add gene and cluster id to the dataframe
    x$gene <- rownames(x)
    x$cluster <- i
    data.frame(x)
})
mgs_df <- do.call(rbind, mgs_fil)
```

```{r ChangeRownames}
#Change to genesymbol whenever possible
rowData(spe)$UniqID <-rownames(spe)
rownames(spe) <- uniquifyFeatureNames(rowData(spe)$ENSEMBL, rowData(spe)$symbol)
data.frame(rowData(spe))
```

```{r }
#Downsample
# split cell indices by identity
idx <- split(seq(ncol(sce)), sce$celltype)
# downsample to at most 20 per identity & subset
# We are using 5 here to speed up the process but set to 75-100 for your real
# life analysis
n_cells <- 75
cs_keep <- lapply(idx, function(i) {
    n <- length(i)
    if (n < n_cells)
        n_cells <- n
    sample(i, n_cells)
})
sce <- sce[, unlist(cs_keep)]
```




```{r }
res <- SPOTlight(
    x = sce,
    y = spe,
    groups = as.character(sce$celltype),
    mgs = mgs_df,
    hvg = rownames(sce)[rowData(sce)$is_hvg],
    weight_id = "mean.AUC",
    group_id = "cluster",
    gene_id = "gene")

head(mat <- res$mat)
# Extract NMF model fit
mod <- res$NMF
```

```{r }
print(plotTopicProfiles(
    x = mod,
    y = sce$celltype,
    facet = FALSE,
    min_prop = 0.01,
    ncol = 1) +
    theme(aspect.ratio = 1))

print(plotTopicProfiles(
    x = mod,
    y = sce$celltype,
    facet = TRUE,
    min_prop = 0.01,
    ncol = 6))

library(NMF)
sign <- basis(mod)
colnames(sign) <- paste0("Topic", seq_len(ncol(sign)))
head(sign)

```


```{r,eval=FALSE}
#look at which genes are the most important for each topic and therefore get an insight into which genes are driving them
basis_spotlight <- data.frame(NMF::basis(mod[[1]]))

colnames(basis_spotlight) <- unique(stringr::str_wrap(mod[[2]], width = 30))

basis_spotlight %>%
  dplyr::arrange(desc(Astro)) %>%
  round(., 5) %>% 
  DT::datatable(., filter = "top")
```


```{r }
#Spatial Correlation Matrix
plotCorrelationMatrix(mat,p.mat = TRUE, hc.order = TRUE,)

#Custom Spatial Correlation Matrix
decon_cor <- cor(mat)
# Compute correlation P-value
p.mat <- corrplot::cor.mtest(mat = mat, conf.level = 0.95)
# Visualize
ggcorrplot::ggcorrplot(
  corr = decon_cor,
  p.mat = p.mat[[1]],
  hc.order = TRUE,
  type = "full",
  insig = "blank",
  lab = TRUE,
  outline.col = "lightgrey",
  method = "square",
  colors = c("#6D9EC1", "white", "#E46726"),
  title = "Predicted cell-cell proportion correlation",
  legend.title = "Correlation\n(Pearson)") +
  ggplot2::theme(
    plot.title = ggplot2::element_text(size = 22, hjust = 0.5, face = "bold"),
    legend.text = ggplot2::element_text(size = 12),
    legend.title = ggplot2::element_text(size = 15),
    axis.text.x = ggplot2::element_text(angle = 90),
    axis.text = ggplot2::element_text(size = 18, vjust = 0.5))

```

### Co-localization
Now that we know which cell types are found within each spot we can make a graph representing spatial interactions where cell types will have stronger edges between them the more often we find them within the same spot.

```{r }
ct <- colnames(mat)
mat[mat < 0.1] <- 0
```

```{r }
#Co-localization
plotInteractions(mat, which = "heatmap", metric = "prop")+ 
  labs(title = "Interaction heatmap proportion", fill = "proportion")#breaks = c(0, 5, 10, 15, 20),

plotInteractions(mat, which = "heatmap", metric = "jaccard")


plotInteractions(mat, which = "network",
    vertex.color = "#cde394",
     vertex.frame.color = "white",
     vertex.label.color = "black",
     vertex.label.family = "Helvetica",
    min_prop = 0.1)
```


```{r }
# Define color palette
# (here we use 'paletteMartin' from the 'colorBlindness' package)
paletteMartin <- c(
    "#000000", "#004949", "#009292", "#ff6db6", "#ffb6db", 
    "#490092", "#006ddb", "#b66dff", "#6db6ff", "#b6dbff", 
    "#920000", "#924900", "#db6d00", "#24ff24", "#ffff6d")

pal <- colorRampPalette(paletteMartin)(length(ct))
names(pal) <- ct

print(
plotSpatialScatterpie(
    x = spe,
    y = mat,
    cell_types = colnames(mat),
    img = FALSE,
    scatterpie_alpha = 1,
    pie_scale = 0.4) +
    scale_fill_manual(
        values = pal,
        breaks = names(pal))
)
#With the image underneath - we are rotating it 90 degrees counterclockwise and mirroring across the horizontal axis to show how to align if the spots don’t overlay the image.
print(
plotSpatialScatterpie(
    x = spe,
    y = mat,
    cell_types = colnames(mat),
    img = TRUE,
    scatterpie_alpha = 1,
    pie_scale = 0.4, 
    degrees = -90,
    axis = "h") +
    scale_fill_manual(
        values = pal,
        breaks = names(pal))
)
```

```{r }
#Residuals
spe$res_ss <- res[[2]][colnames(spe)]
xy <- spatialCoords(spe)
spe$x <- xy[, 1]
spe$y <- xy[, 2]
ggcells(spe, aes(x, y, color = res_ss)) +
    geom_point() +
    scale_color_viridis_c() +
    coord_fixed() +
    theme_bw()
```

```{r }
library(viridis)

for(i in colnames(mat)){
  print(plotSpots(spe, annotate = mat[,colnames(mat)==i],size = 1.7)+labs(title=paste0("Spatial coordinates ",i, " proportion"))+
  scale_color_viridis(option="magma")
        )
}

```

## Save the dataset

```{r SaveData}
saveRDS(mat, file =paste0(path,'/output/Spatial_knee_SC.rds'))
saveRDS(spe, file =paste0(path,'/output/Spatial_knee.rds'))
```


## Deconvolution Annotation_L2

```{r }
sce <- readRDS(file = paste0(path,'/output/07_sce_SC_Merge.rds'))
#exclude some celltypes
sce <- sce[,sce$celltype != "Neutrophil"]


#Create Annotation
sce$SpatialAnno <- sce$Annotation_L2
unique(sce$SpatialAnno)

sce$SpatialAnno[sce$Annotation_L2 %in% c("CytotoxCD4 TC","Memory TC","Naive TC","CytotoxCD8 TC","NKTcell")] <- "T cell"
sce$SpatialAnno[sce$Annotation_L2 %in% c("TREM2+ MP","FOLR2+ MP")] <- "Tissue-resident Macrophage"
sce$SpatialAnno[sce$Annotation_L2 %in% c("S100A12+ MP","CD48+ MP")] <- "Macrophage"
sce$SpatialAnno[sce$Annotation_L2 %in% c("cDC","pDC")] <- "Dendritic cell"
sce$SpatialAnno[sce$Annotation_L2 %in% c("Arteriolar EC","Venular EC","Capillary EC")] <- "Endothelial"
#sce$SpatialAnno[sce$Annotation_L2 %in% c("PRG4+ FIB","CHI3L2+ FIB")] <- "Lining Fibroblast"
#sce$SpatialAnno[sce$Annotation_L2 %in% c("POSTN+ FIB","CXCL12+ FIB","MFAP5+ FIB")] <- "Sub-lining Fibroblast"
unique(sce$SpatialAnno)

```


```{r }
mgs <- scoreMarkers(sce,groups=sce$SpatialAnno, subset.row = rownames(sce)[rowData(sce)$hvg.sc])
mgs_fil <- lapply(names(mgs), function(i) {
    x <- mgs[[i]]
    # Filter and keep relevant marker genes, those with AUC > 0.8
    x <- x[x$mean.AUC > 0.5, ]
    # Sort the genes from highest to lowest weight
    x <- x[order(x$mean.AUC, decreasing = TRUE), ]
    # Add gene and cluster id to the dataframe
    x$gene <- rownames(x)
    x$cluster <- i
    data.frame(x)
})
mgs_df <- do.call(rbind, mgs_fil)
```

```{r ChangeRownames}
#Change to genesymbol whenever possible
rowData(spe)$UniqID <-rownames(spe)
rownames(spe) <- uniquifyFeatureNames(rowData(spe)$ENSEMBL, rowData(spe)$symbol)
data.frame(rowData(spe))
```

```{r }
#Downsample
# split cell indices by identity
idx <- split(seq(ncol(sce)), sce$SpatialAnno)
# downsample to at most 20 per identity & subset
# We are using 5 here to speed up the process but set to 75-100 for your real
# life analysis
n_cells <- 100
cs_keep <- lapply(idx, function(i) {
    n <- length(i)
    if (n < n_cells)
        n_cells <- n
    sample(i, n_cells)
})
sce <- sce[, unlist(cs_keep)]
```




```{r }
res <- SPOTlight(
    x = sce,
    y = spe,
    groups = as.character(sce$SpatialAnno),
    mgs = mgs_df,
    hvg = rownames(sce)[rowData(sce)$is_hvg_sc],
    weight_id = "mean.AUC",
    group_id = "cluster",
    gene_id = "gene")

head(mat <- res$mat)
# Extract NMF model fit
mod <- res$NMF
```

```{r }
print(plotTopicProfiles(
    x = mod,
    y = sce$SpatialAnno,
    facet = FALSE,
    min_prop = 0.01,
    ncol = 1) +
    theme(aspect.ratio = 1))

print(plotTopicProfiles(
    x = mod,
    y = sce$SpatialAnno,
    facet = TRUE,
    min_prop = 0.01,
    ncol = 6))

library(NMF)
sign <- basis(mod)
colnames(sign) <- paste0("Topic", seq_len(ncol(sign)))
head(sign)

```


```{r,eval=FALSE}
#look at which genes are the most important for each topic and therefore get an insight into which genes are driving them
basis_spotlight <- data.frame(NMF::basis(mod[[1]]))

colnames(basis_spotlight) <- unique(stringr::str_wrap(mod[[2]], width = 30))

basis_spotlight %>%
  dplyr::arrange(desc(Astro)) %>%
  round(., 5) %>% 
  DT::datatable(., filter = "top")
```


```{r }
#Spatial Correlation Matrix
plotCorrelationMatrix(mat)
```

### Co-localization
Now that we know which cell types are found within each spot we can make a graph representing spatial interactions where cell types will have stronger edges between them the more often we find them within the same spot.

```{r }
ct <- colnames(mat)
mat[mat < 0.1] <- 0
```

```{r }
#Co-localization
plotInteractions(mat, which = "heatmap", metric = "prop")+ 
  labs(title = "Interaction heatmap proportion", fill = "proportion") #breaks = c(0, 5, 10, 15, 20),

plotInteractions(mat, which = "heatmap", metric = "jaccard")+ 
  labs(title = "Interaction heatmap", fill = "jaccard")+ 


plotInteractions(mat, which = "network",
    vertex.color = "#cde394",
     vertex.frame.color = "white",
     vertex.label.color = "black",
     vertex.label.family = "Helvetica",
    min_prop = 0.1)
```


```{r }
# Define color palette
# (here we use 'paletteMartin' from the 'colorBlindness' package)
paletteMartin <- c(
    "#000000", "#004949", "#009292", "#ff6db6", "#ffb6db", 
    "#490092", "#006ddb", "#b66dff", "#6db6ff", "#b6dbff", 
    "#920000", "#924900", "#db6d00", "#24ff24", "#ffff6d")

pal <- colorRampPalette(paletteMartin)(length(ct))
names(pal) <- ct

print(
plotSpatialScatterpie(
    x = spe,
    y = mat,
    cell_types = colnames(mat),
    img = FALSE,
    scatterpie_alpha = 1,
    pie_scale = 0.4) +
    scale_fill_manual(
        values = pal,
        breaks = names(pal))
)
#With the image underneath - we are rotating it 90 degrees counterclockwise and mirroring across the horizontal axis to show how to align if the spots don’t overlay the image.
print(
plotSpatialScatterpie(
    x = spe,
    y = mat,
    cell_types = colnames(mat),
    img = TRUE,
    scatterpie_alpha = 1,
    pie_scale = 0.4, 
    degrees = -90,
    axis = "h") +
    scale_fill_manual(
        values = pal,
        breaks = names(pal))
)
```

```{r }
#Residuals
spe$res_ss <- res[[2]][colnames(spe)]
xy <- spatialCoords(spe)
spe$x <- xy[, 1]
spe$y <- xy[, 2]
ggcells(spe, aes(x, y, color = res_ss)) +
    geom_point() +
    scale_color_viridis_c() +
    coord_fixed() +
    theme_bw()
```

```{r }
library(viridis)

for(i in colnames(mat)){
  print(plotSpots(spe, annotate = mat[,colnames(mat)==i],size = 1.7)+labs(title=paste0("Spatial coordinates ",i, " proportion"))+
  scale_color_viridis(option="magma")
        )
}

```

## Save the dataset

```{r SaveData}

saveRDS(spe, file =paste0(path,'/output/Spatial_knee_SC.rds'))
saveRDS(res, file =paste0(path,'/output/Spatial_knee_SC.rds'))
```
